### 2021-09-14 ProjectPart1.Rmd
Temp Delete Data

---
title: "Project Part 1 - Data Description and Questions to Answer"
author: "Ryan Grunert"
date: "9/10/2021"
output: word_document
---

### Data Description

The data I will be using for the MADA class project is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017. The data was collected in a collaboration between Dr. Michael Yabsley's lab (of which I am a part) and Dr. Sonia Hernandez's lab here at UGA. I received the raw data from Dr. Yabsley in an agreement to analyze the data over the course of the semester in exchange for using the data for the MADA project. The parasite the data focuses on is the protozoan *Haemoproteus*, which causes a chronic infection in not only White Ibises, but many other birds as well. *Haemoproteus* is a vector-borne disease, it is transmitted by *Culicoides* biting midges when they feed on the birds and causes long-lasting parasitemia.

The data is split into two sections; 2010-2014 and 2015-2017. For 2010-2014, there are 269 observations (each of which represents a sample taken from different individual birds) and 20 variables. For 2015-2017, there are 401 observations and 29 variables.

The two sections differ in the number of variables recorded for each observation.

### 2021-09-17 products/manuscript/Manuscript.Rmd
Completed Part 1 for Project
Completed the first part for the MADA project. The raw data files are in the raw data folder. The main text and glimpses of the data in code are in the "Manuscript.Rmd" file in the products folder.

---
title: "MADA Project - Ibis Blood Parasites from Florida"
subtitle: ""
author: Ryan Grunert
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---
```{r}
#Loading required packages
library(dplyr) #for data processing
library(here) #to set paths
library(tidyverse) #all required data manipulation packages
library(xlsx) #to read in excel spreadsheets
```

# Introduction

## General Background Information
The data I will be using for this project is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017. The parasite the data focuses on is the protozoan *Haemoproteus*, which causes a chronic blood infection in White Ibises and many other birds. *Haemoproteus* is a vector-borne disease, it is transmitted by *Culicoides* biting midges when they feed on the birds and causes long-lasting parasitemia. Adult healthy birds tolerate haemosporidian infections well, but young birds and birds outside their normal range that are naive to the pathogen are more likely to develop clinical disease. Haemosporidian infections usually cause long-lasting chronic infections in natural hosts, and can cause reduced reproductive success, host fitness, and increased stress and disease susceptibility.

Here is a link to a previously published study concerning this data -> https://www.sciencedirect.com/science/article/pii/S221322441730072X

## Description of data and data source
The data was collected in a collaboration between Dr. Michael Yabsley's lab (of which I am a part) and Dr. Sonia Hernandez's lab here at UGA. I received the raw data from Dr. Yabsley in an agreement to analyze the data over the course of the semester in exchange for using the data for the MADA project.

The data is organized in excel spreadsheets. The two main spreadsheets (Ibis_Blood_ParasiteData_2017 and IbisFieldDataOct19_2017) hold the majority of the data, and two supplemental spreadsheets from other collaborators that were or currently working on the project contain additional variables that will be combined with the main dataset in R for additional analyses. Each observation in the data is an individual White Ibis that was captured and sampled, and each variable is a different descriptor of either the Ibis itself, or the site/environment it was captured in.

Two datasheets have overlapping information (Ibis_Blood_ParasiteData_2017 and IbisFieldDataOct19_2017), but both will be combined to have a complete dataset for analysis. The first contains parasitemia data, while the second contains field data. The 100IbisID_UrbanGradient datasheet will be used to assist with the analysis concerning *Haemoproteus* presence and habitat type. The SalmonellaFieldSerotypes datasheet may be used to compare the presence of both *Haemoproteus* and *Salmonella* in the Ibis, although that may not be needed considering the difference in biology and epidemiology of the two pathogens.

```{r}
#Setting the path to the Ibis blood parasite spreadsheet
data_locationIbisB <- here::here("data","raw_data","Ibis_Blood_ParasiteData_2017.xlsx")

#Loads the raw Ibis blood data excel sheet from 2015-2017
rdIbisBlood15_17 <- read.xlsx(data_locationIbisB, sheetName = "2015-2017")

#Loads the raw Ibis blood data excel sheet from 2010-2014
rdIbisBlood10_14 <- read.xlsx(data_locationIbisB, sheetName = "2010-2014")

glimpse(rdIbisBlood15_17)
glimpse(rdIbisBlood10_14)

```

The main variables I'm interested in for these two sheets above are "Haemoproteus.parasitemia.", and the first "X" column is the sample/bird ID. The 2010-14 set is missing more data, but characteristics of the birds from that group may be used.

```{r}
#Setting the path to the Ibis field data spreadsheet
data_locationIbisF <- here::here("data","raw_data","IbisFieldDataOct19_2017.xlsx")

#Loads the raw Ibis field data excel sheet
rdIbisField <- read.xlsx(data_locationIbisF, sheetName = "All capture data")

glimpse(rdIbisField)

```

This dataset is the most organized sheet, but columns of interest will be extracted from this sheet as well as the others to create new clean data frames for analysis.

```{r}
#Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
data_locationIbisUr <- here::here("data","raw_data","100IbisID_UrbanGradient.xlsx")

#Loads the raw Ibis Urban Gradient supplemental data sheet
UrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")

glimpse(UrbanIbis)

```

These samples from this datasheet will be used in conjunction with the two main datasheets in order to create an "urbanization gradient" for the samples that are located in either rural or urban areas. The gradient will most likely range from least ("1", rural) to most ("4", urban) urban for example, but further work and planning needs to be done first in order to figure out the specifics of the analysis.

```{r}
#Setting the path to the second Ibis supplemental excel sheet (Salmonella)
data_locationIbisSal <- here::here("data","raw_data","SalmonellaFieldSerotypesAug27_2018.xlsx")

#Loads the raw Ibis Urban Gradient supplemental data sheet
SalIbis <- read.xlsx(data_locationIbisSal, sheetName = "ALL Positives")

glimpse(SalIbis)
```

This sheet details the Ibises that tested positive for Salmonella, but this data may not be used because of the differences in pathogenesis and epidemiology of the two diseases.

## Questions/Hypotheses to be addressed
The questions I plan to address are whether the amounts of *Haemoproteus* parasitemia in the various Ibises are significantly different when focusing on age, sex, season, location, or habitat type (Urban and Rural). Does the amount of parasitemia in the birds decrease, increase or remain the same over time as they get older? Is parasitemia significantly higher in birds sampled in urban or rural environments? These are the main variables I will be looking at for this project, but I may analyze others as the project continues. Not all of the variables were collected for every Ibis sampled, and the data still needs cleaning to accurately detail the sample size for each analysis.

## Expected Analyses
I think that I will have to conduct multiple ANOVA's during the analysis process. Linear regression models may apply here too for some variables, but I'm not entirely sure. I would like to explore the possibility of using the location data for each sample to conduct a geographical analysis with a map as the output. I will be continuing to contact those involved with the study as well to see what analyses they want to be conducted and what questions they want answered.

# Methods and Results

## Data aquisition

## Data import and cleaning

## Exploratory analysis

## Full analysis

# Discussion

## Summary and Interpretation

## Strengths and Limitations

## Conclusions

# References

# Summary/Abstract
_Write a summary of your project._

### 2021-10-07 products/manuscript/Manuscript.Rmd
Completed Project Part 2 10/7/21
Completed up to Part 2 of the MADA project. Data cleaning and exploratory analysis have been finished, along with all organization of the folders and files. Updated READMEs. Will return in later parts to data cleaning or exploratory analysis as needed.

@@
 ```{r}
-#Loading required packages
-library(dplyr) #for data processing
+#Loading required packages #for data processing
 library(here) #to set paths
@@
 library(xlsx) #to read in excel spreadsheets
+library(janitor) #To assist in cleaning up dates and tidy data
+library(lubridate) #To extract times from raw data
 ```
@@
 
-glimpse(rdIbisBlood15_17)
-glimpse(rdIbisBlood10_14)
+str(rdIbisBlood15_17)
+str(rdIbisBlood10_14)
 
@@
 
-glimpse(rdIbisField)
-
+str(rdIbisField)
 ```
@@
 #Loads the raw Ibis Urban Gradient supplemental data sheet
-UrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")
+rdUrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")
 
-glimpse(UrbanIbis)
+str(rdUrbanIbis)
 
@@
 
-glimpse(SalIbis)
+str(SalIbis)
 ```
@@
 
+Is there a difference in HaeParasit between Ibis of different ages?
+- Dependent variable (outcome): HaeParasit
+- Independent variable (predictor): Age
+
+ Is there a difference in HaeParasit between Ibis of different sex?
+- Dependent variable (outcome): HaeParasit
+- Independent variable (predictor): Sex
+
+Is there a different in HaeParasit in Ibis sampled at different sites?
+ - Dependent variable (outcome): HaeParasit
+ - Independent variable (predictor): Site
+
+ Is there a difference in HaeParasit between Ibis sampled in different habitat types?
+ - Dependent variable (outcome): HaeParasit
+ - Independent variable (predictor): HabType
+
+ Is there a difference in HaeParasit between Ibis sampled at different times?
+ - Dependent variable (outcome): HaeParasit
+ - Independent variables (predictors): Date, Season
+
 ## Expected Analyses
 I think that I will have to conduct multiple ANOVA's during the analysis process.
-Linear regression models may apply here too for some variables, but I'm not entirely sure.
-I would like to explore the possibility of using the location data for each sample to conduct a geographical analysis with a map as the output.
+Linear and logistic regression models apply here too for some variables.I would like to explore the possibility of using the location data for each sample to conduct a geographical analysis with a map as the output.
 I will be continuing to contact those involved with the study as well to see what analyses they want to be conducted and what questions they want answered.
 
-# Methods and Results
+## Data Cleaning
+The data has already been imported, so now we move onto the cleaning stage.
+The following code will clean, process, and save the raw Ibis blood datasets from 2010-2014 and 2015-2017.
+This section is contained in the "ProcessingScriptIbisBloodData.R" file.
+```{r}
+##############################Cleaning the rdIbisBlood10_14 data###########################
+IbisBlood10_14 <- rename(rdIbisBlood10_14, ID = X.) #renames the ID column
 
-## Data aquisition
+#Extracts just the columns of interest
+IbisBlood10_14 <- select(IbisBlood10_14, ID, Site.ID, Date, PCR.sex, Age..adult.juvenile., 
+                         Total.RBC, Haemoproteus.parasitemia.)
 
-## Data import and cleaning
+#Renames the columns of interest
+IbisBlood10_14 <- rename(IbisBlood10_14, Site = Site.ID, Sex = PCR.sex, Age = Age..adult.juvenile.,
+                         TotalRBC = Total.RBC, HaeParasit = Haemoproteus.parasitemia.)
+
+#Change "adult" and "juvenile" in the Age column to "A" and "J"
+IbisBlood10_14$Age[IbisBlood10_14$Age == "adult"] <- "A"
+IbisBlood10_14$Age[IbisBlood10_14$Age == "Adult"] <- "A"
+IbisBlood10_14$Age[IbisBlood10_14$Age == "juvenile"] <- "J"
+IbisBlood10_14$Age[IbisBlood10_14$Age == "SA"] <- "A"
+#Now the Ibises that had age recorded are referred to as "A" for adult or "J" for juvenile
+
+#Changing the DIV/0 values in HaeParasit to NA and change the column to numerical values
+IbisBlood10_14$HaeParasit <- as.numeric(IbisBlood10_14$HaeParasit)  # Divide-by-zero errors are now NA
+
+#Reformatting the date values in the Date column
+IbisBlood10_14$Date <- excel_numeric_to_date(IbisBlood10_14$Date)
+#After this step, the values that just state the year are labelled with "1905", may need to go back or exclude from analysis
+
+#Abbreviating the site names to make analysis easier
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Royal Palms"] <- "RP"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Everglades"] <- "E"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Dreher Park"] <- "DRP"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Indian Creek"] <- "ICP"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Indian Creek "] <- "ICP"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Indian Creek Park"] <- "ICP"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Juno Beach"] <- "JB"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Lion Country Safari"] <- "LCS"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Lake Worth"] <- "LW"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Prosperity Oaks"] <- "PO"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Promenade Plaza"] <- "PP"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "San Mateo"] <- "SM"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "Solid Waste Authority"] <- "SWA"
+IbisBlood10_14$Site[IbisBlood10_14$Site == "West Palm Beach Zoo"] <- "WPBZ"
+
+#Reformatting the mislabeled dates. These are set to Jan 1st of the year they were
+#recorded on the data sheet, these entries only had the year stated, not month and day.
+IbisBlood10_14$Date[IbisBlood10_14$Date == "1905-07-02"] <- "2010-01-01"
+IbisBlood10_14$Date[IbisBlood10_14$Date == "1905-07-03"] <- "2011-01-01"
+IbisBlood10_14$Date[IbisBlood10_14$Date == "1905-07-04"] <- "2012-01-01"
+
+#The bottom of the dataset has a couple rows of just missing data for all variables,
+#This code filters the dataset so that if any rows are present that has at least
+#one variable be NOT NA, R keeps it
+IbisBlood10_14 <- filter(IbisBlood10_14, if_any(everything(), ~ !is.na(.)))
+
+#This completes the cleaning phase for IbisBlood10_14
+#------------------------------------------------------------------------------------#
+
+############################Cleaning the rdIbisBlood15_17 Data#########################
+
+IbisBlood15_17<- rename(rdIbisBlood15_17, ID= X.) #renames the ID column
+
+####Extracting only the columns relevant for analysis
+IbisBlood15_17 <- select(IbisBlood15_17,ID,Site.Name,Date,PCR.sex,Age,Season,Habitat.type,
+                         Total.RBC,X..Haemoproteus,Haemoproteus.parasitemia.,F1,F2,F3,
+                         Avg.RBC.FOV,Ibis..,Total.mass..g.,Mass.bag..g.,Mass.bird..g.,Body.condition.score..1.5.,
+                         Ectoparasite.score..1.5.,Culmen.length..mm.,Wing.chord.length..mm.,
+                         Tarsus.length..mm.,Tarsus.width..mm.)
+
+####Renaming the columns 
+IbisBlood15_17 <- rename(IbisBlood15_17,Site = Site.Name,Sex = PCR.sex,HabType = Habitat.type,TotalRBC = Total.RBC,NumHae = X..Haemoproteus,
+                         HaeParasit = Haemoproteus.parasitemia.,AvgRBC = Avg.RBC.FOV,
+                         IbisNum = Ibis..,TotalMassG = Total.mass..g.,BagMassG = Mass.bag..g.,
+                         BirdMassG = Mass.bird..g.,BodyCondScore = Body.condition.score..1.5.,
+                         EctoParasitScore = Ectoparasite.score..1.5.,CulmenLmm = Culmen.length..mm.,
+                         WingChordLmm = Wing.chord.length..mm.,TarsusLmm = Tarsus.length..mm.,
+                         TarsusWmm = Tarsus.width..mm.)
+
+####Site column
+#Ibis that have sites matched with the 10_14 dataset will have the same labels
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Juno Beach"] <- "JB"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Indian Creek Park"] <- "ICP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Dubois park"] <- "DUP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Dreher Park"] <- "DRP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Lion Country Safari"] <- "LCS"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Dubois Park"] <- "DUP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "J.W. Corbett Wildlife Management Area"] <- "JWC"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Gaines Park"] <- "GP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Solid Waste Authority"] <- "SWA"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Fisheating creek"] <- "FC"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Kitching Creek"] <- "KC"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Loxahatchee NE"] <- "LOXNE"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "DuBois Park"] <- "DUP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "TetraTech"] <- "TT"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Green Cay"] <- "GC"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Dreher park"] <- "DRP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "CROW Wildlife Rehabilitation Center"] <- "CROW"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Lake Worth"] <- "LW"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Royal Palm"] <- "RP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "JW Corbett"] <- "JWC"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Busch Wildlife Sanctuary"] <- "BWS"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Loxahatchee Slough"] <- "LOXS"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Indian Creek"] <- "ICP"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Cat House (Bonnie's, Richard Lane)"] <- "CAT"
+IbisBlood15_17$Site[IbisBlood15_17$Site == "Palm Beach Zoo"] <- "WPBZ"
+
+####Date column
+#Changing the date column to numeric values to allow date extraction
+IbisBlood15_17$Date <- as.numeric(IbisBlood15_17$Date) 
+##Reformatting the date values in the Date column
+IbisBlood15_17$Date <- excel_numeric_to_date(IbisBlood15_17$Date)
+
+####Sex column
+IbisBlood15_17$Sex[IbisBlood15_17$Sex == "not bled"] <- NA
+#Two values have missing data because the birds were not bled to determine sex
+IbisBlood15_17$Sex[IbisBlood15_17$Sex == "M?"] <- NA
+IbisBlood15_17$Sex[IbisBlood15_17$Sex == "F?"] <- NA
+IbisBlood15_17$Sex[IbisBlood15_17$Sex == "F "] <- "F"
+#Two values have question marks for unsure sex determination, removing those values
+
+####Age column
+#When an ibis reaches 4 years old, their plumage becomes all white. Any Ibis
+#that was sampled with brown plumage is considered a juvenile. Since Haemoproteus is
+#a chronic disease that mostly likely accumulates over time, long-lived birds should
+#theoretically have more parasitemia. Could not be true though and may come back to
+#look at the individual years in the juvenile birds
+IbisBlood15_17$Age[IbisBlood15_17$Age == "Juvenile"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "Adult"] <- "A"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "2nd/3rd year"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "3rd year"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "1st year"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "adult"] <- "A"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "2nd year"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "NA"] <- NA
+IbisBlood15_17$Age[IbisBlood15_17$Age == "2nd Year"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "Late third year"] <- "J"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "ADULT"] <- "A"
+IbisBlood15_17$Age[IbisBlood15_17$Age == "Adult bc"] <- "A" #Not sure what bc stands for
+
+#Season and HabType are fine the way they are
+
+####TotalRBC
+IbisBlood15_17$TotalRBC <- as.numeric(IbisBlood15_17$TotalRBC)
+#DIV/0 errors are now NA in this column
+
+#NumHae column is fine
+
+####HaeParasit
+IbisBlood15_17$HaeParasit <- as.numeric(IbisBlood15_17$HaeParasit)
+
+####-999 values
+IbisBlood15_17[IbisBlood15_17 == -999] <- NA
+
+####Removing the last couple empty rows
+IbisBlood15_17 <- IbisBlood15_17[-c(401, 402, 403, 404, 405, 406),]
+
+#This completes the cleaning phase for IbisBlood15_17
+#------------------------------------------------------------------------------------#
+
+###############################Saving the Cleaned Datasets###########################
+
+#Saving the Ibis Blood 10-14 dataset as an RDS in the processed data folder
+save_data_location <- here::here("data", "processed_data", "processedIbisBlood10_14.rds")
+saveRDS(IbisBlood10_14, file = save_data_location)
+
+#Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
+save_data_location2 <- here::here("data", "processed_data", "processedIbisBlood15_17.rds")
+saveRDS(IbisBlood15_17, file = save_data_location2)
+
+```
+
+This next section will clean, process, and save the raw Ibis Field dataset. This section is contained in the "ProcessingScriptIbisFieldData.R" file.
+
+```{r}
+######################### #Cleaning the rdIbisField Data###############################
+
+IbisField <- rename(rdIbisField, ID = ID..NSFSITE...) #renames the ID column
+
+#Selecting the columns of interest to this analysis from the datasheet
+IbisField <- select(IbisField,ID,Site.Name,Latitude,Longitude,Date,PCR.Sex,Age,Season,
+                    Habitat.type,Ibis.number,Mass.bird..g.,Body.condition.score..1.5.,
+                    Ectoparasite.score..1.5.,Culmen.length..mm.,Wing.chord.length..mm.,
+                    Tarsus.length..mm.,Tarsus.width..mm.,Habituation.score.of.flock..1.5.)
+
+#Renaming the columns of interest, also matching the other cleaned datasets
+IbisField <- rename(IbisField,Site = Site.Name,Lat = Latitude,Long = Longitude,Sex = PCR.Sex,
+                    HabType = Habitat.type,IbisNum = Ibis.number,BirdMassG = Mass.bird..g.,
+                    BodyCondScore = Body.condition.score..1.5.,EctoParasitScore = Ectoparasite.score..1.5.,
+                    CulmenLmm = Culmen.length..mm.,WingChordLmm = Wing.chord.length..mm.,
+                    TarsusLmm = Tarsus.length..mm.,TarsusWmm = Tarsus.width..mm.,HabFlockScore = Habituation.score.of.flock..1.5.)
+
+#This dataset has "-999" values which mean they are intentionally left blank,
+#Birds with an ID number of "-999" represent sites that didn't catch any birds,
+#but had weather data collected. This code deletes those observations.
+IbisField <- IbisField[!(IbisField$ID == "-999"),]
+
+####Site
+#These site location abbreviations are the same across datasets
+IbisField$Site[IbisField$Site == "Busch Wildlife Sanctuary"] <- "BWS"
+IbisField$Site[IbisField$Site == "Cat House (Bonnie's, Richard Lane)"] <- "CAT"
+IbisField$Site[IbisField$Site == "Dubois Park"] <- "DUP"
+IbisField$Site[IbisField$Site == "DuBois Park"] <- "DUP"
+IbisField$Site[IbisField$Site == "Dreher Park"] <- "DRP"
+IbisField$Site[IbisField$Site == "Fisheating Creek"] <- "FC"
+IbisField$Site[IbisField$Site == "Green Cay"] <- "GC"
+IbisField$Site[IbisField$Site == "Gaines Park"] <- "GP"
+IbisField$Site[IbisField$Site == "Indian Creek"] <- "ICP"
+IbisField$Site[IbisField$Site == "Juno Beach"] <- "JB"
+IbisField$Site[IbisField$Site == "Juno Beach "] <- "JB"
+IbisField$Site[IbisField$Site == "J.W. Corbett Wildlife Management Area"] <- "JWC"
+IbisField$Site[IbisField$Site == "Kitching Creek"] <- "KC"
+IbisField$Site[IbisField$Site == "Lion Country Safari"] <- "LCS"
+IbisField$Site[IbisField$Site == "Lake Worth"] <- "LW"
+IbisField$Site[IbisField$Site == "Loxahatchee Slough"] <- "LOXS"
+IbisField$Site[IbisField$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
+IbisField$Site[IbisField$Site == "Loxahatchee/LILA Cell B2"] <- "LOXB2"
+IbisField$Site[IbisField$Site == "Loxahatchee NE "] <- "LOXNE"
+IbisField$Site[IbisField$Site == "Palm Beach Zoo"] <- "WPBZ"
+IbisField$Site[IbisField$Site == "Royal Palm"] <- "RP"
+IbisField$Site[IbisField$Site == "CROW Wildlife Rehabilitation Center"] <- "CROW"
+IbisField$Site[IbisField$Site == "Solid Waste Authority"] <- "SWA"
+IbisField$Site[IbisField$Site == "TetraTech"] <- "TT"
+
+####Sex Column
+#This replaces the -999 and FAIL values with NA. The IbisBlood datasheets carry more sex data for the birds
+IbisField$Sex[IbisField$Sex == "FAIL"] <- NA
+
+####Age Column
+IbisField$Age[IbisField$Age == "Adult"] <- "A"
+IbisField$Age[IbisField$Age == "3rd year"] <- "J"
+IbisField$Age[IbisField$Age == "2nd year"] <- "J"
+IbisField$Age[IbisField$Age == "1st year"] <- "J"
+IbisField$Age[IbisField$Age == "NA"] <- NA
+IbisField$Age[IbisField$Age == "Adult "] <- "A"
+IbisField$Age[IbisField$Age == "2nd year "] <- "J"
+IbisField$Age[IbisField$Age == "Juvenile"] <- "J"
+IbisField$Age[IbisField$Age == "2nd Year"] <- "J"
+
+#Replacing all -999 values in the dataset with NA
+IbisField[IbisField == -999] <- NA
+
+#Replacing NA text strings with regular NA values in multiple columns
+IbisField$BodyCondScore[IbisField$BodyCondScore == "NA"] <- NA
+IbisField$HabFlockScore[IbisField$HabFlockScore == "NA"] <- NA
+IbisField$CulmenLmm[IbisField$CulmenLmm == "NA"] <- NA
+IbisField$WingChordLmm[IbisField$WingChordLmm == "NA"] <- NA
+IbisField$TarsusLmm[IbisField$TarsusLmm == "NA"] <- NA
+IbisField$TarsusWmm[IbisField$TarsusWmm == "NA"] <- NA
+
+#--------------------------------------------------------------------------------------#
+
+###############################Saving the Cleaned Dataset###############################
+
+save_data_location3 <- here::here("data", "processed_data", "processedIbisFielddata.rds")
+saveRDS(IbisField, file = save_data_location3)
+```
+
+This cleaning code section will clean, process, and save the Ibis urbanization dataset. This section is contained in the "ProcessingScriptIbisUrban.R" file.
+
+```{r}
+
+##############################Cleaning the rdUrbanIbis data###########################
+#Selects only the columns of interest in the dataset
+IbisUrban <- select(rdUrbanIbis, Name:Serotype)
+
+#Renames the columns of interest
+IbisUrban <- rename(IbisUrban,ID = Name,Date = Collection.Date,HgPPM = mg.kg.Hg..PPM.,
+                    Site = Site.Name,UrbanPercent = Site...Urbanized)
+
+#Change the site names to their abbreviations
+IbisUrban$Site[IbisUrban$Site == "Juno Beach"] <- "JB"
+IbisUrban$Site[IbisUrban$Site == "Indian Creek"] <- "ICP"
+IbisUrban$Site[IbisUrban$Site == "Dubois Park"] <- "DUP"
+IbisUrban$Site[IbisUrban$Site == "Dreher Park"] <- "DRP"
+IbisUrban$Site[IbisUrban$Site == "Lion Country Safari"] <- "LCS"
+IbisUrban$Site[IbisUrban$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
+IbisUrban$Site[IbisUrban$Site == "Solid Waste Authority "] <- "SWA"
+IbisUrban$Site[IbisUrban$Site == "Gaines Park"] <- "GP"
+IbisUrban$Site[IbisUrban$Site == "Kitching Creek"] <- "KC"
+IbisUrban$Site[IbisUrban$Site == "Kitching Creek "] <- "KC"
+IbisUrban$Site[IbisUrban$Site == "Loxahatchee NE"] <- "LOXNE"
+IbisUrban$Site[IbisUrban$Site == "TetraTech"] <- "TT"
+IbisUrban$Site[IbisUrban$Site == "Loxahatchee NE "] <- "LOXNE"
+IbisUrban$Site[IbisUrban$Site == "Green Cay"] <- "GC"
+IbisUrban$Site[IbisUrban$Site == "J.W. Corbett Wildlife Management Area"] <- "JWC"
+IbisUrban$Site[IbisUrban$Site == "Loxahatchee Slough"] <- "LOXS"
+IbisUrban$Site[IbisUrban$Site == "Lake Worth"] <- "LW"
+
+#--------------------------------------------------------------------------------------#
+
+################################Saving the UrbanIbis Data#############################
+#Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
+save_data_location3 <- here::here("data", "processed_data", "processedIbisUrban.rds")
+saveRDS(IbisUrban, file = save_data_location3)
+
+#--------------------------------------------------------------------------------------#
+```
+
+This last cleaning section combines the data sheets into one to use for analysis.
+The urbanization gradient has not been created as of yet, but a meeting this month with those involved with the project will help shed some light on the exact specifics for analysis.
+This code is located in the "ExploratoryAnalysisScript.R" file.
+
+```{r}
+###############################Completing Combination Data Frames#########################
+#Outer joining IbisBlood10_14 and IbisBlood15_17, keeping all values of the 10_14 dataset
+IbisBlood10_17 <- merge(IbisBlood10_14, IbisBlood15_17, by = c("ID", "Site", "Date", "Sex", "Age",
+                                                              "TotalRBC", "HaeParasit"), all=TRUE)
+#There is no overlap in the ID numbers between the two, so no recapture is assumed.
+
+#This is a left join, this data frame below only contains birds that were in the IbisBlood10_17
+#dataset (all with parasitemia data), with additional information added by the IbisField data. Ibis
+#from the Field data that weren't tested for parasitemia or weren't included on Blood dataset
+#are not included in this dataframe
+
+#IbisBlood10_17(with Field)_(Parasitemia data)
+IbisC <- left_join(IbisBlood10_17, IbisField, by = c("ID", "Site", "Date", "Sex", "Age", "Season", "HabType",
+                                                         "IbisNum", "BirdMassG", "BodyCondScore", "EctoParasitScore",
+                                                         "CulmenLmm", "WingChordLmm", "TarsusLmm", "TarsusWmm"))
+
+#This section classifies the habitat type for the sites without a habitat type.
+IbisC <- IbisC %>%
+        mutate(HabType = ifelse(Site == "ICP", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "JB", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "RP", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "LCS", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "LW", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "SWA", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "E", "Natural", HabType)) %>%
+        mutate(HabType = ifelse(Site == "WPBZ", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "PO", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "PP", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "DRP", "Urban", HabType)) %>%
+        mutate(HabType = ifelse(Site == "SM", "Urban", HabType))
+
+##This section is adding in the seasons for observations that have collection dates but no season.
+
+#Creating intervals for adding seasons based on the date column
+int1 <- interval(ymd("2012-12-01"), ymd("2012-12-31"))
+int2 <- interval(ymd("2013-06-01"), ymd("2013-08-31"))
+int3 <- interval(ymd("2013-12-01"), ymd("2013-12-31"))
+int4 <- interval(ymd("2014-02-01"), ymd("2014-03-31"))
+
+#Adding seasons to the ibis without them stated based on the date the sameple was collected
+IbisC <- IbisC %>%
+         mutate(Season = ifelse(Date %within% int1, "Fall 2012", Season)) %>%
+         mutate(Season = ifelse(Date %within% int2, "Summer 2013", Season)) %>%
+         mutate(Season = ifelse(Date %within% int3, "Fall 2013", Season)) %>%
+         mutate(Season = ifelse(Date %within% int4, "Spring 2014", Season))
+
+#Adding a column for HaeParasit Presence/Absence data
+IbisC <- IbisC %>%
+    mutate(HaeParasitPA = as.integer(HaeParasit > 0 & !is.na(HaeParasit)))
+
+#Adding a column for transformed HaeParasit data on a logarithm base 10 scale,
+#then removing the -Inf values due to there being no parasitemia in those birds. This column
+#then only contains logarithm transformed parasitemia data for birds with a positive value
+#of parasitemia.
+IbisC <- IbisC %>%
+    mutate(HaeParasitLog10 = log10(HaeParasit)) %>%
+    mutate(HaeParasitLog10 = ifelse(HaeParasitLog10 == "-Inf", NA, HaeParasitLog10))
+
+#Changing the BodyCondScore to an ordinal scale
+IbisC <- IbisC %>%
+      mutate(BodyCondScore = recode_factor(BodyCondScore, '1' = 'Emaciated',
+                                           '2' = 'Underweight', '3' = 'Ideal',
+                                           '4' = 'Overweight', '5' = 'Obese',
+                                           .ordered = TRUE))
+
+#Changing the class of some variables for analysis. From character to numeric.
+C2N <- c(16,17,18,21,22,23,24)
+
+IbisC[ , C2N] <- apply(IbisC[ , C2N], 2,
+                       function(x) as.numeric(as.character(x)))
+
+```
+
+This concludes the data cleaning section, we are now ready for the exploratory analysis!
 
@@
 
-## Full analysis
+### Tables
 
-# Discussion
+To start the exploratory analysis, we're going to create some quick tables to get an idea of the metrics for some of the columns of interest.
 
-## Summary and Interpretation
+```{r}
+IbisC %>%
+  count(Sex, wt = HaeParasitPA)
+```
 
-## Strengths and Limitations
+Looking at presence/absence data for haemoproteus parasitemia, there were about double the number of female ibises with active infection compared to the males.
+78 of the ibises sampled were negative for haemoproteus infection.
 
-## Conclusions
+```{r}
+IbisC %>%
+  count(Age, wt = HaeParasitPA)
+```
 
-# References
+Here we can see that there was about 2.7 times more adults sampled with active infection than juvenile ibises.
+59 of the sampled ibises with active infection either couldn't be aged or age was not recorded.
 
-# Summary/Abstract
-_Write a summary of your project._
+```{r}
+IbisC %>%
+  count(Age, Sex, wt = HaeParasitPA)
+```
+
+Combining age and sex, we can see that adult females had the highest number of active infections. Categories with missing numbers are prevalent.
+
+```{r}
+IbisC %>%
+  count(Site, wt = HaeParasitPA)
+```
+Looking at site, SWA, ICP, and DRP were the sites with the highest number of sampled ibises with active infection.
+
+```{r}
+IbisC %>%
+  count(Season, wt = HaeParasitPA)
+```
+Looking at season sampled, Spring 2016 had the highest number of sampled ibises with active infection, followed by Summer 2013, Spring 2017, and Spring 2014.
+
+```{r}
+IbisC %>%
+  count(BodyCondScore, Age, wt = HaeParasitPA)
+```
+Looking at body condition score, the majority of the birds sampled with active infection were an ideal weight or underweight.
+This lets us know that an active infection didn't affect body condition too harshly on these birds.
+
+### Figures
+
+Now we can move onto creating some figures for the data to see any patterns visually.
+
+Figure 1: Haemoproteus parasitemia from sampled infected Ibises from 2010-2017 by habitat type
+```{r}
+ExpFigure1 <-ggplot(IbisC) +
+  geom_point(aes(x = Date, y = HaeParasitLog10, color = HabType)) +
+  ggtitle("Haemoproteus parasitemia from sampled infected Ibises from 2010-2017 by habitat type") +
+  xlab("Date") +
+  ylab("Haemoproteus Parasitemia (Log base 10)") +
+  scale_color_discrete("Habitat Type", na.translate = F) +
+  theme(plot.title = element_text(hjust = 0.5, size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        legend.title = element_text(size = 10))
+
+Figure_file1 <- here("results", "ExploratoryAnalysis", "ExpFigure1.png")
+ggsave(filename = Figure_file1,plot = ExpFigure1)
+
+ExpFigure1
+```
+
+This scatterplot graphs the sample dates chronologically against the parasitemia measured from the sampled ibis on a transformed log base 10 scale.
+Habitat type is depicted by color of the dots.
+Overall, there is a general narrowing of the data over time with a slight negative correlation.
+There isn't a clear difference between natural and urban habitat types when it comes to amount of parasitemia.
+
+Figure 2: Haemoproteus parasitemia from adult and juvenile sampled Ibises
+```{r}
+ExpFigure2 <- IbisC %>%
+  drop_na(Age) %>%
+  ggplot() +
+  geom_violin(aes(x = Age, y = HaeParasit)) +
+  ggtitle("Haemoproteus parasitemia from adult and juvenile sampled Ibises") +
+  xlab("Age") +
+  ylab("Haemoproteus Parasitemia") +
+  scale_y_continuous(limits = c(0, 0.1)) +
+  theme(plot.title = element_text(hjust = 0.5, size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        legend.title = element_text(size = 10))
+
+Figure_file2 <- here("results", "ExploratoryAnalysis", "ExpFigure2.png")
+ggsave(filename = Figure_file2, plot = ExpFigure2)
+
+ExpFigure2
+```
+
+This violin plot graphs the haemoproteus parasitemia distribution for the adult and juvenile Ibises sampled.
+The majority of the parasitemia values are on the lower end for both adults and juveniles.
+The distribution for adults reaches higher than juveniles.
+The overall structure for the adult distribution is higher and skinnier, while juveniles have more of their parasitemias lower in value.
+
+Figure 3: Haemoproteus parasitemia from sampled Ibises by sex
+```{r}
+ExpFigure3 <- IbisC %>%
+  drop_na(Sex) %>%
+  ggplot() +
+  geom_boxplot(aes(x = Sex, y = HaeParasit)) +
+  scale_y_continuous(limits = c(0, 0.5)) +
+  ggtitle("Haemoproteus parasitemia from sampled Ibises by sex") +
+  xlab("Sex") +
+  ylab("Haemoproteus Parasitemia") +
+  theme(plot.title = element_text(hjust = 0.5, size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        legend.title = element_text(size = 10))
+
+Figure_file3 <- here("results", "ExploratoryAnalysis", "ExpFigure3.png")
+ggsave(filename = Figure_file3, plot = ExpFigure3)
+
+ExpFigure3
+```
+
+The haemoproteus parasitemia distributions between sexes is extremely similar.
+The median value for males is slightly higher than females.
+The majority of the measured parasitemias are between 0.0 and 0.1.
+
+Figure 4: Haemoproteus parasitemia by total female and male Ibis mass
+```{r}
+ExpFigure4 <- IbisC %>%
+  drop_na(BirdMassG, Sex) %>%
+  ggplot() +
+  geom_point(aes(x = BirdMassG, y = HaeParasitLog10, color = Sex)) +
+  theme(axis.text.x = element_text(angle = 90)) +
+  scale_x_continuous(limits = c(130, 1240)) +
+  ggtitle("Haemoproteus parasitemia by total female and male Ibis mass") +
+  xlab("Total mass of Ibis (g)") +
+  ylab("Haemoproteus Parasitemia (Log base 10)") +
+  theme(plot.title = element_text(hjust = 0.5, size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        legend.title = element_text(size = 10))
+
+Figure_file4 <- here("results", "ExploratoryAnalysis", "ExpFigure4.png")
+ggsave(filename = Figure_file4, plot = ExpFigure4)
+
+ExpFigure4
+```
+
+Looking at the plot, male birds seem to weigh more than females. However, when comparing the parasitemia between females and males, there isn't a clear difference between the two.
+
+Figure 5: Haemoproteus parasitemia by female and male Ibis mass and body condition score
+```{r}
+ExpFigure5 <- IbisC %>%
+  drop_na(BirdMassG, BodyCondScore, Sex) %>%
+  ggplot() +
+  geom_point(aes(x = BirdMassG, y = HaeParasitLog10, color = BodyCondScore)) +
+  facet_wrap(~ Sex) +
+  ggtitle("Haemoproteus parasitemia by female and male Ibis mass and body condition score") +
+  xlab("Total mass of Ibis (g)") +
+  ylab("Haemoproteus Parasitemia (Log base 10)") +
+  theme(axis.text.x = element_text(angle = 90),
+        plot.title = element_text(hjust = 0.5, size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        legend.title = element_text(size = 10)) +
+  scale_x_continuous(limits = c(130, 1240)) +
+  scale_color_discrete("Body Condition", na.translate = F)
+
+Figure_file5 <- here("results", "ExploratoryAnalysis", "ExpFigure5.png")
+ggsave(filename = Figure_file5, plot = ExpFigure5)
+
+ExpFigure5
+```
+
+Looking at this plot, there are more underweight female ibises with higher parasitemias compared to males. This could be due to the higher number of sampled females though.
+
+Figure 6: Haemoproteus parasitemia by sample site and habitat type
+```{r}
+ExpFigure6 <- IbisC %>%
+  drop_na(HabType, Site) %>%
+  ggplot() +
+  geom_point(aes(x = Site, y = log10(HaeParasit), color = HabType)) +
+  theme(axis.text.x = element_text(angle = 90),
+        plot.title = element_text(hjust = 0.5, size = 10),
+        axis.title.x = element_text(size = 10),
+        axis.title.y = element_text(size = 10),
+        legend.title = element_text(size = 10)) +
+  ggtitle("Haemoproteus parasitemia by sample site and habitat type") +
+  xlab("Sample Site") +
+  ylab("Haemoproteus Parasitemia (Log base 10)") +
+  scale_color_discrete("Habitat Type", na.translate = F)
+
+Figure_file6 <- here("results", "ExploratoryAnalysis", "ExpFigure6.png")
+ggsave(filename = Figure_file6, plot = ExpFigure6)
+
+ExpFigure6
+```
+
+Looking at this plot broadly, it seems that the sites with a natural habitat have a slightly lower average parasitemia than the sites with an urban habitat.
+
+Overall, there are no huge patterns or clear differences between covariates and variates so far.
+Statistical analyses may pick up on any significant differences later on, but I will return to the exploratory analysis as I continue with the project.
+I will also be meeting with those involved with the project to discuss as well.

### 2021-10-07 code/analysis_code/ExploratoryAnalysisScript.R

@@ -1,16 +1,16 @@
 ###############################
 # Exploratory Analysis Script
 #
-#this script loads the processed and cleaned data and performs an intitial 
+#this script loads the processed and cleaned data and performs an initial 
 #exploratory analysis on the data. This script will also include a section on variates/covariates
-#and what questions to answer. Will combine into an Rmarkdown document for better visuals
+#and what questions to answer. Will combine into an Rmarkdown document for better visuals.
 
 
 ###############################Loading all required packages###########################
 
 library(here) #to set paths
 library(tidyverse) #all required data manipulation packages
-
+library(lubridate) #work with date/time data
 #--------------------------------------------------------------------------------------#
 
 ###############################Loading all datasets####################################
@@ -28,23 +28,207 @@ IbisUrban <- readRDS(data_location4)
 
 #--------------------------------------------------------------------------------------#
 
-####Variates/Covariates
+###############################Completing Combination Data Frames#########################
+#Outer joining IbisBlood10_14 and IbisBlood15_17, keeping all values of the 10_14 dataset
+IbisBlood10_17 <- merge(IbisBlood10_14, IbisBlood15_17, by = c("ID", "Site", "Date", "Sex", "Age",
+                                                              "TotalRBC", "HaeParasit"), all=TRUE)
+#There is no overlap in the ID numbers between the two, so no recapture is assumed.
+
+#This is a left join, this data frame below only contains birds that were in the IbisBlood10_17
+#dataset (all with parasitemia data), with additional information added by the IbisField data. Ibis
+#from the Field data that weren't tested for parasitemia or weren't included on Blood dataset
+#are not included in this dataframe
+
+#IbisBlood10_17(with Field)_(Parasitemia data)
+IbisC <- left_join(IbisBlood10_17, IbisField, by = c("ID", "Site", "Date", "Sex", "Age", "Season", "HabType",
[…]

### 2021-10-07 code/processing_code/ProcessingScriptIbisBloodData.R

@@ -93,7 +93,7 @@ IbisBlood15_17 <- select(IbisBlood15_17,ID,Site.Name,Date,PCR.sex,Age,Season,Hab
                          Total.RBC,X..Haemoproteus,Haemoproteus.parasitemia.,F1,F2,F3,
                          Avg.RBC.FOV,Ibis..,Total.mass..g.,Mass.bag..g.,Mass.bird..g.,Body.condition.score..1.5.,
                          Ectoparasite.score..1.5.,Culmen.length..mm.,Wing.chord.length..mm.,
-                         Tarsus.length..mm.,Tarsus.width..mm.,Volume.serology.bleed.B.and.C..ul..after.bka.and.cort.)
+                         Tarsus.length..mm.,Tarsus.width..mm.)
                          
 ####Renaming the columns 
 IbisBlood15_17 <- rename(IbisBlood15_17,Site = Site.Name,Sex = PCR.sex,HabType = Habitat.type,TotalRBC = Total.RBC,NumHae = X..Haemoproteus,
@@ -102,7 +102,7 @@ IbisBlood15_17 <- rename(IbisBlood15_17,Site = Site.Name,Sex = PCR.sex,HabType =
                          BirdMassG = Mass.bird..g.,BodyCondScore = Body.condition.score..1.5.,
                          EctoParasitScore = Ectoparasite.score..1.5.,CulmenLmm = Culmen.length..mm.,
                          WingChordLmm = Wing.chord.length..mm.,TarsusLmm = Tarsus.length..mm.,
-                         TarsusWmm = Tarsus.width..mm.,SeroVolBleed = Volume.serology.bleed.B.and.C..ul..after.bka.and.cort.)
+                         TarsusWmm = Tarsus.width..mm.)
 
 ####Site column
 #Ibis that have sites matched with the 10_14 dataset will have the same labels
@@ -144,6 +144,7 @@ IbisBlood15_17$Sex[IbisBlood15_17$Sex == "not bled"] <- NA
 #Two values have missing data because the birds were not bled to determine sex
 IbisBlood15_17$Sex[IbisBlood15_17$Sex == "M?"] <- NA
 IbisBlood15_17$Sex[IbisBlood15_17$Sex == "F?"] <- NA
+IbisBlood15_17$Sex[IbisBlood15_17$Sex == "F "] <- "F"
 #Two values have question marks for unsure sex determination, removing those values
 
 ####Age column
[…]

### 2021-10-18 products/manuscript/Manuscript.Rmd
Altered Structure of Project Repo 10/18/21
Completely changed the structure of the overall project repo. Manuscript still needs altering to accommodate the changes.

@@
 #Setting the path to the Ibis blood parasite spreadsheet
-data_locationIbisB <- here::here("data","raw_data","Ibis_Blood_ParasiteData_2017.xlsx")
+data_locationIbisB <- here::here("raw_data","Ibis_Blood_ParasiteData_2017.xlsx")
 
@@
 #Setting the path to the Ibis field data spreadsheet
-data_locationIbisF <- here::here("data","raw_data","IbisFieldDataOct19_2017.xlsx")
+data_locationIbisF <- here::here("raw_data","IbisFieldDataOct19_2017.xlsx")
 
@@
 #Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
-data_locationIbisUr <- here::here("data","raw_data","100IbisID_UrbanGradient.xlsx")
+data_locationIbisUr <- here::here("raw_data","100IbisID_UrbanGradient.xlsx")
 
@@
 #Setting the path to the second Ibis supplemental excel sheet (Salmonella)
-data_locationIbisSal <- here::here("data","raw_data","SalmonellaFieldSerotypesAug27_2018.xlsx")
+data_locationIbisSal <- here::here("raw_data","SalmonellaFieldSerotypesAug27_2018.xlsx")
 
@@
 #Saving the Ibis Blood 10-14 dataset as an RDS in the processed data folder
-save_data_location <- here::here("data", "processed_data", "processedIbisBlood10_14.rds")
+save_data_location <- here::here("processed_data", "processedIbisBlood10_14.rds")
 saveRDS(IbisBlood10_14, file = save_data_location)
@@
 #Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
-save_data_location2 <- here::here("data", "processed_data", "processedIbisBlood15_17.rds")
+save_data_location2 <- here::here("processed_data", "processedIbisBlood15_17.rds")
 saveRDS(IbisBlood15_17, file = save_data_location2)

### 2021-10-18 analysis_scripts/ExploratoryAnalysisScript.R

@@ -16,10 +16,10 @@ library(lubridate) #work with date/time data
 ###############################Loading all datasets####################################
 #This code loads all of the processed and cleaned Ibis data
 
-data_location1 <- here::here("data","processed_data","processedIbisBlood10_14.rds")
-data_location2 <- here::here("data","processed_data","processedIbisBlood15_17.rds")
-data_location3 <- here::here("data","processed_data","processedIbisFielddata.rds")
-data_location4 <- here::here("data","processed_data","processedIbisUrban.rds")
+data_location1 <- here::here("processed_data","processedIbisBlood10_14.rds")
+data_location2 <- here::here("processed_data","processedIbisBlood15_17.rds")
+data_location3 <- here::here("processed_data","processedIbisFielddata.rds")
+data_location4 <- here::here("processed_data","processedIbisUrban.rds")
 
 IbisBlood10_14 <- readRDS(data_location1)
 IbisBlood15_17 <- readRDS(data_location2)
@@ -136,7 +136,7 @@ ExpFigure1 <-ggplot(IbisC) +
         axis.title.y = element_text(size = 10),
         legend.title = element_text(size = 10))
 
-Figure_file1 <- here("results", "ExploratoryAnalysis", "ExpFigure1.png")
+Figure_file1 <- here("results", "Figures", "ExpFigure1.png")
 ggsave(filename = Figure_file1,plot = ExpFigure1)
 
 #########Plot 2: Violin plot of haemoproteus parasitemia vs adult and juvenile
@@ -154,7 +154,7 @@ ExpFigure2 <- IbisC %>%
         axis.title.y = element_text(size = 10),
         legend.title = element_text(size = 10))
 
-Figure_file2 <- here("results", "ExploratoryAnalysis", "ExpFigure2.png")
+Figure_file2 <- here("results", "Figures", "ExpFigure2.png")
 ggsave(filename = Figure_file2, plot = ExpFigure2)
 
 #########Plot 3: Boxplot of haemoproteus parasitemia vs sex
@@ -172,7 +172,7 @@ ExpFigure3 <- IbisC %>%
         axis.title.y = element_text(size = 10),
         legend.title = element_text(size = 10))
 
-Figure_file3 <- here("results", "ExploratoryAnalysis", "ExpFigure3.png")
[…]

### 2021-10-18 code/processing_code/ProcessingScriptIbisFieldData.R

@@ -1,104 +1,0 @@
-###############################
-# Processing Script 2 - ProcessingScriptIbisFieldData.R
-#This processing script cleans and processes the IbisFieldDataOct19_2017 datasheet, which contains
-#field data regarding the sampled ibis for Haemoproteus parasitemia
-
-###############################Loading all required packages###########################
-library(here) #to set paths
-library(tidyverse) #all required data manipulation packages
-library(xlsx) #to read in excel spreadsheets
-library(lubridate) #to extract times from raw data
-library(janitor) #to help clean up dates and tidy data
-
-#--------------------------------------------------------------------------------------#
-
-##################################Inputting Raw Data###################################
-#Setting the path to the Ibis field data spreadsheet
-data_locationIbisF <- here::here("data","raw_data","IbisFieldDataOct19_2017.xlsx")
-#Loads the raw Ibis field data excel sheet
-rdIbisField <- read.xlsx(data_locationIbisF, sheetName = "All capture data")
-
-#--------------------------------------------------------------------------------------#
-
-######################### #Cleaning the rdIbisField Data###############################
-
-IbisField <- rename(rdIbisField, ID = ID..NSFSITE...) #renames the ID column
-
-#Selecting the columns of interest to this analysis from the datasheet
-IbisField <- select(IbisField,ID,Site.Name,Latitude,Longitude,Date,PCR.Sex,Age,Season,
-                    Habitat.type,Ibis.number,Mass.bird..g.,Body.condition.score..1.5.,
-                    Ectoparasite.score..1.5.,Culmen.length..mm.,Wing.chord.length..mm.,
-                    Tarsus.length..mm.,Tarsus.width..mm.,Habituation.score.of.flock..1.5.)
-
-#Renaming the columns of interest, also matching the other cleaned datasets
-IbisField <- rename(IbisField,Site = Site.Name,Lat = Latitude,Long = Longitude,Sex = PCR.Sex,
-                    HabType = Habitat.type,IbisNum = Ibis.number,BirdMassG = Mass.bird..g.,
[…]

### 2021-10-18 code/processing_code/ProcessingScriptIbisUrban.R

@@ -1,63 +1,0 @@
-###################
-#Processing Script 3 - 100IbisID_UrbanGradient
-#This script loads and cleans the Urbanization dataset for the Ibis parasitemia data.
-
-###############################Loading all required packages###########################
-library(here) #to set paths
-library(tidyverse) #all required data manipulation packages
-library(xlsx) #to read in excel spreadsheets
-library(lubridate) #to extract times from raw data
-library(janitor) #to help clean up dates and tidy data
-
-#--------------------------------------------------------------------------------------#
-
-##################################Inputting Raw Data###################################
-#Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
-data_locationIbisUr <- here::here("data","raw_data","100IbisID_UrbanGradient.xlsx")
-#Loads the raw Ibis Urban Gradient supplemental data sheet
-rdUrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")
-
-#--------------------------------------------------------------------------------------#
-
-##############################Cleaning the rdUrbanIbis data###########################
-#Selects only the columns of interest in the dataset
-IbisUrban <- select(rdUrbanIbis, Name:Serotype)
-
-#Renames the columns of interest
-IbisUrban <- rename(IbisUrban,ID = Name,Date = Collection.Date,HgPPM = mg.kg.Hg..PPM.,
-                    Site = Site.Name,UrbanPercent = Site...Urbanized)
-
-#Change the site names to their abbreviations
-IbisUrban$Site[IbisUrban$Site == "Juno Beach"] <- "JB"
-IbisUrban$Site[IbisUrban$Site == "Indian Creek"] <- "ICP"
-IbisUrban$Site[IbisUrban$Site == "Dubois Park"] <- "DUP"
-IbisUrban$Site[IbisUrban$Site == "Dreher Park"] <- "DRP"
-IbisUrban$Site[IbisUrban$Site == "Lion Country Safari"] <- "LCS"
-IbisUrban$Site[IbisUrban$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
-IbisUrban$Site[IbisUrban$Site == "Solid Waste Authority "] <- "SWA"
-IbisUrban$Site[IbisUrban$Site == "Gaines Park"] <- "GP"
[…]

### 2021-10-18 processing_script/ProcessingScript.R

@@ -1,9 +1,6 @@
-###############################
-# Processing Script 1 - ProcessingScriptIbisBloodData.R
-#This script is to load, clean, and process the main parasitemia dataset containing
-#Ibis parasitemia data. Other scripts are used for cleaning and processing the
-#Ibis urban and Salmonella datasheets, and Ibis field datasheet.
-
+###################
+#Processing Script
+#This script loads and cleans raw data for the Ibis data analysis process. It then saves processed data for further analysis.
 
 ###############################Loading all required packages###########################
 library(here) #to set paths
@@ -14,17 +11,22 @@ library(janitor) #to help clean up dates and tidy data
 
 #--------------------------------------------------------------------------------------#
 
-
 ##################################Inputting Raw Data###################################
 #Setting the path to the Ibis blood parasite spreadsheet
-data_locationIbisB <- here::here("data","raw_data","Ibis_Blood_ParasiteData_2017.xlsx")
+data_locationIbisB <- here::here("raw_data","Ibis_Blood_ParasiteData_2017.xlsx")
+#Setting the path to the Ibis field data spreadsheet
+data_locationIbisF <- here::here("raw_data","IbisFieldDataOct19_2017.xlsx")
+#Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
+data_locationIbisUr <- here::here("raw_data","100IbisID_UrbanGradient.xlsx")
 #Loads the raw Ibis blood data excel sheet from 2015-2017
 rdIbisBlood15_17 <- read.xlsx(data_locationIbisB, sheetName = "2015-2017")
 #Loads the raw Ibis blood data excel sheet from 2010-2014
 rdIbisBlood10_14 <- read.xlsx(data_locationIbisB, sheetName = "2010-2014")
-
+#Loads the raw Ibis field data excel sheet
+rdIbisField <- read.xlsx(data_locationIbisF, sheetName = "All capture data")
+#Loads the raw Ibis Urban Gradient supplemental data sheet
+rdUrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")
[…]

### 2021-10-29 products/manuscript/Manuscript.Rmd
Completed Project Part 3 10/29/21
Completed part 3 of the MADA project. Statistical analysis scripts contains code to create 10 models, and save them. Analyses afterwards will most likely focus on the presence absence data, and work on tweaking the GLM models already in place.

@@
 
-save_data_location3 <- here::here("data", "processed_data", "processedIbisFielddata.rds")
+save_data_location3 <- here::here("processed_data", "processedIbisFielddata.rds")
 saveRDS(IbisField, file = save_data_location3)
@@
 #Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
-save_data_location3 <- here::here("data", "processed_data", "processedIbisUrban.rds")
+save_data_location3 <- here::here("processed_data", "processedIbisUrban.rds")
 saveRDS(IbisUrban, file = save_data_location3)
@@
 
-Figure_file1 <- here("results", "ExploratoryAnalysis", "ExpFigure1.png")
+Figure_file1 <- here("results", "Figures", "ExpFigure1.png")
 ggsave(filename = Figure_file1,plot = ExpFigure1)
@@
 
-Figure_file2 <- here("results", "ExploratoryAnalysis", "ExpFigure2.png")
+Figure_file2 <- here("results", "Figures", "ExpFigure2.png")
 ggsave(filename = Figure_file2, plot = ExpFigure2)
@@
 
-Figure_file3 <- here("results", "ExploratoryAnalysis", "ExpFigure3.png")
+Figure_file3 <- here("results", "Figures", "ExpFigure3.png")
 ggsave(filename = Figure_file3, plot = ExpFigure3)
@@
 
-Figure_file4 <- here("results", "ExploratoryAnalysis", "ExpFigure4.png")
+Figure_file4 <- here("results", "Figures", "ExpFigure4.png")
 ggsave(filename = Figure_file4, plot = ExpFigure4)
@@
 
-Figure_file5 <- here("results", "ExploratoryAnalysis", "ExpFigure5.png")
+Figure_file5 <- here("results", "Figures", "ExpFigure5.png")
 ggsave(filename = Figure_file5, plot = ExpFigure5)
@@
 
-Figure_file6 <- here("results", "ExploratoryAnalysis", "ExpFigure6.png")
+Figure_file6 <- here("results", "Figures", "ExpFigure6.png")
 ggsave(filename = Figure_file6, plot = ExpFigure6)
@@
 I will also be meeting with those involved with the project to discuss as well.
+
+## Statistical Analyses
+
+### Presence/Absence Haemoproteus Parasitemia Data
+
+This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data just looking at the presence and absence data, as opposed to the amount of parasitemia per sampled bird.
+
+The following models were setup and tested through a 10-fold cross-validation.
+The training/testing data proportions were 80/20.
+10 folds were taken on 80% of the observations, and the results are depicted here.
+I am currently struggling with applying the models generated by CV to the testing data.
+
+```{r}
+data_locationPA1 <- here::here("results", "Tables", "Model1MetricsPA.Rda")
+table1 <- readRDS(data_locationPA1)
+table1
+```
+This univariate classification GLM looked at the effect of habitat type on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+
+```{r}
+data_locationPA2 <- here::here("results", "Tables", "Model2MetricsPA.Rda")
+table2 <- readRDS(data_locationPA2)
+table2
+```
+This univariate classification GLM looked at the effect of sex of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+
+```{r}
+data_locationPA3 <- here::here("results", "Tables", "Model3MetricsPA.Rda")
+table3 <- readRDS(data_locationPA3)
+table3
+```
+This univariate classification GLM looked at the effect of age of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+
+```{r}
+data_locationPA4 <- here::here("results", "Tables", "Model4MetricsPA.Rda")
+table4 <- readRDS(data_locationPA4)
+table4
+```
+This univariate classification GLM looked at the effect of sampling site on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+
+```{r}
+data_locationPA5 <- here::here("results", "Tables", "Model5MetricsPA.Rda")
+table5 <- readRDS(data_locationPA5)
+table5
+```
+This univariate classification GLM looked at the effect of sampling date on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+
+```{r}
+data_locationPA6 <- here::here("results", "Tables", "Model6MetricsPA.Rda")
+table6 <- readRDS(data_locationPA6)
+table6
+```
+This univariate classification GLM looked at the effect of sampling season on the presence of Haemoproteus in infected birds.
+The results were unsatisfactory, however, these results were better than the other univariate models so far.
+
+```{r}
+data_locationPA7 <- here::here("results", "Tables", "Model7MetricsPA.Rda")
+table7 <- readRDS(data_locationPA7)
+table7
+```
+This multivariate classification GLM looked at the effect of age and sex on the presence of Haemoproteus in infected birds.
+The results were unsatisfactory, this model scored about the same as the previous univariate models for age and sex.
+
+```{r}
+data_locationPA8 <- here::here("results", "Tables", "Model8MetricsPA.Rda")
+table8 <- readRDS(data_locationPA8)
+table8
+```
+This multivariate classification GLM looked at the effect of sampling date and sampling season on the presence of Haemoproteus in infected birds.
+The results here had the highest performance of all the other models so far.
+One way to interpret this model would be that as the sampling effort continued from 2010-2017, the passage of time and season had a significant effect on the presence of Haemoproteus in Ibis.
+However, the performance of this model isn't high enough to successfully conclude with that explanation.
+
+### Regression Models for the Amount of Haemoproteus Parasitemia
+
+This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data looking at the amount of measured parasitemia per sampled bird (on a logarithm base 10 scale), as opposed to presence/absence data.
+
+The following models were setup and tested through a 10-fold cross-validation.
+The training/testing data proportions were 80/20.
+10 folds were taken on 80% of the observations, and the results are depicted here.
+I am currently struggling with applying the models generated by CV to the testing data.
+
+#### Random Forest Model
+
+The following are the results from a random forest model run on all the predictors against the Haemoproteus parasitemia (log base 10) data.
+The missing data values were imputed by K nearest neighbors.
+The model was run on all 10 folds from the training dataset.
+```{r}
+data_locationRF <- here::here("results", "Tables", "RandomForestMetrics.Rda")
+table9 <- readRDS(data_locationRF)
+table9
+```
+This random forest model looked at the effect of all predictors on the amount of Haemoproteus in infected birds.
+The results were unsatisfactory, this model had a relatively high RMSE and low R squared value.
+
+#### Linear Regression Model
+
+The following are the results from a linear regression model run on habitat type against the Haemoproteus parasitemia (log base 10) data.
+
+```{r}
+data_locationLM1 <- here::here("results", "Tables", "LinMetrics1.Rda")
+table10 <- readRDS(data_locationLM1)
+table10
+```
+This linear regression model looked at the effect of habitat type on the amount of Haemoproteus in infected birds.
+The results were unsatisfactory, the RMSE is high and the R squared value is very low.
+
+#### Conclusions so far
+
+This concludes the statistical analysis for part 3.
+Overall, the models did not perform well, especially those looking at the amount of Haemoproteus in infected birds.
+The presence/absence data has more potential for conclusions to be made based on generated models.
+The next steps would be to fine-tune the classification models for the presence/absence data and create a random forest model.
+
+Current problems that have arose in this process include imputing missing values to run models (as there is a large amount of missing data in this dataset), and visualizing the tidymodels results on a graph.

### 2021-10-29 analysis_scripts/ExploratoryAnalysisScript.R

@@ -103,10 +103,15 @@ C2N <- c(16,17,18,21,22,23,24)
 IbisC[ , C2N] <- apply(IbisC[ , C2N], 2,
                        function(x) as.numeric(as.character(x)))
 
+##############################Saving the Complete Dataset#########################
+
+save_data_location <- here::here("processed_data", "CompleteIbisDataset.rds")
+
+saveRDS(IbisC, file = save_data_location)
+
 ###############################Exploratory Tables#################################
 IbisC %>%
 count(Sex, wt = HaeParasitPA)
-
 
 IbisC %>%
   count(Age, wt = HaeParasitPA)
@@ -231,4 +236,8 @@ ExpFigure6 <- IbisC %>%
 
 Figure_file6 <- here("results", "Figures", "ExpFigure6.png")
 ggsave(filename = Figure_file6, plot = ExpFigure6)
+
+
+
+
 

### 2021-10-29 analysis_scripts/StatisticalAnalysisModels.R

@@ -1,0 +1,283 @@
+###############################################################################
+# Statistical Analysis Script
+#
+#This script loads the processed data and runs the statistical analyses for the project.
+#This includes creating and saving all models, figures, and tables associated with the analysis.
+
+###############################Loading all required packages###########################
+library(here) #to set paths
+library(tidyverse) #all required data manipulation packages
+library(lubridate) #work with date/time data
+library(tidymodels) #to run the statistical analyses
+library(ranger) #implementation of random forests
+
+set.seed(144)
+#--------------------------------------------------------------------------------------#
+
+###########################Loading Data for Analysis##############################
+data_location5 <- here::here("processed_data","CompleteIbisDataset.rds")
+
+IbisC <- readRDS(data_location5)
+#--------------------------------------------------------------------------------------#
+############################################################################################################
+################################Presence/Absence Data for Haemoproetus Parasitemia##########################
+############################################################################################################
+
+########This section is setting up models for categorical binomial presence absence data
+#####Setting up Data Frame
+#This extracts the columns we need for the analysis
+IbisPAdf <- IbisC[c("HaeParasitPA", "Sex", "Age", "Site", "Season",
+                    "BodyCondScore", "Date", "HabType", "BirdMassG")] #Extracts columns of interest
+
+#Changing columns to factors for use in making models
+IbisPAdf$HaeParasitPA <- as.factor(IbisPAdf$HaeParasitPA)
+IbisPAdf$Sex <- as.factor(IbisPAdf$Sex)
+IbisPAdf$Age <- as.factor(IbisPAdf$Age)
+IbisPAdf$Site <- as.factor(IbisPAdf$Site)
+IbisPAdf$Season <- as.factor(IbisPAdf$Season)
[…]

### 2021-11-11 products/manuscript/Manuscript.Rmd
Finished Statistical Analysis Part 4 11/11/21
Finished the statistical analysis for the project, organization and manuscript writing left.

@@
 ---
-title: "MADA Project - Ibis Blood Parasites from Florida"
+title: "Prevalence of *Haemoproteus plataleae* in American White Ibis from Southern Florida"
 subtitle: ""
@@
 ## General Background Information
-The data I will be using for this project is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017.
+The data this analysis concerns is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017.
 The parasite the data focuses on is the protozoan *Haemoproteus*, which causes a chronic blood infection in White Ibises and many other birds.
@@
 
-Here is a link to a previously published study concerning this data -> https://www.sciencedirect.com/science/article/pii/S221322441730072X
+The purpose of this analysis is to determine whether the presence of *Haemoproteus* was increasing over time in Ibis in southern Florida, and if so, determine any demographic trends among the Ibis or temporal or spatial trends among the system at large.
+Here we investigate whether age, sex, sampling site, capture season, date of capture, body condition, and/or mass of captured Ibis make significant differences on the overall presence of *Haemoproteus* in the birds.
 
-## Description of data and data source
-The data was collected in a collaboration between Dr.
-Michael Yabsley's lab (of which I am a part) and Dr.
-Sonia Hernandez's lab here at UGA.
-I received the raw data from Dr.
-Yabsley in an agreement to analyze the data over the course of the semester in exchange for using the data for the MADA project.
+# Methods
 
-The data is organized in excel spreadsheets.
-The two main spreadsheets (Ibis_Blood_ParasiteData_2017 and IbisFieldDataOct19_2017) hold the majority of the data, and two supplemental spreadsheets from other collaborators that were or currently working on the project contain additional variables that will be combined with the main dataset in R for additional analyses.
-Each observation in the data is an individual White Ibis that was captured and sampled, and each variable is a different descriptor of either the Ibis itself, or the site/environment it was captured in.
+## Data Collection
 
-Two datasheets have overlapping information (Ibis_Blood_ParasiteData_2017 and IbisFieldDataOct19_2017), but both will be combined to have a complete dataset for analysis.
-The first contains parasitemia data, while the second contains field data.
-The 100IbisID_UrbanGradient datasheet will be used to assist with the analysis concerning *Haemoproteus* presence and habitat type.
-The SalmonellaFieldSerotypes datasheet may be used to compare the presence of both *Haemoproteus* and *Salmonella* in the Ibis, although that may not be needed considering the difference in biology and epidemiology of the two pathogens.
-
-```{r}
-#Setting the path to the Ibis blood parasite spreadsheet
-data_locationIbisB <- here::here("raw_data","Ibis_Blood_ParasiteData_2017.xlsx")
-
-#Loads the raw Ibis blood data excel sheet from 2015-2017
-rdIbisBlood15_17 <- read.xlsx(data_locationIbisB, sheetName = "2015-2017")
-
-#Loads the raw Ibis blood data excel sheet from 2010-2014
-rdIbisBlood10_14 <- read.xlsx(data_locationIbisB, sheetName = "2010-2014")
-
-str(rdIbisBlood15_17)
-str(rdIbisBlood10_14)
-
-```
-
-The main variables I'm interested in for these two sheets above are "Haemoproteus.parasitemia.", and the first "X" column is the sample/bird ID.
-The 2010-14 set is missing more data, but characteristics of the birds from that group may be used.
-
-```{r}
-#Setting the path to the Ibis field data spreadsheet
-data_locationIbisF <- here::here("raw_data","IbisFieldDataOct19_2017.xlsx")
-
-#Loads the raw Ibis field data excel sheet
-rdIbisField <- read.xlsx(data_locationIbisF, sheetName = "All capture data")
-
-str(rdIbisField)
-```
-
-This dataset is the most organized sheet, but columns of interest will be extracted from this sheet as well as the others to create new clean data frames for analysis.
-
-```{r}
-#Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
-data_locationIbisUr <- here::here("raw_data","100IbisID_UrbanGradient.xlsx")
-
-#Loads the raw Ibis Urban Gradient supplemental data sheet
-rdUrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")
-
-str(rdUrbanIbis)
-
-```
-
-These samples from this datasheet will be used in conjunction with the two main datasheets in order to create an "urbanization gradient" for the samples that are located in either rural or urban areas.
-The gradient will most likely range from least ("1", rural) to most ("4", urban) urban for example, but further work and planning needs to be done first in order to figure out the specifics of the analysis.
-
-```{r}
-#Setting the path to the second Ibis supplemental excel sheet (Salmonella)
-data_locationIbisSal <- here::here("raw_data","SalmonellaFieldSerotypesAug27_2018.xlsx")
-
-#Loads the raw Ibis Urban Gradient supplemental data sheet
-SalIbis <- read.xlsx(data_locationIbisSal, sheetName = "ALL Positives")
-
-str(SalIbis)
-```
-
-This sheet details the Ibises that tested positive for Salmonella, but this data may not be used because of the differences in pathogenesis and epidemiology of the two diseases.
-
-## Questions/Hypotheses to be addressed
-The questions I plan to address are whether the amounts of *Haemoproteus* parasitemia in the various Ibises are significantly different when focusing on age, sex, season, location, or habitat type (Urban and Rural).
-Does the amount of parasitemia in the birds decrease, increase or remain the same over time as they get older?
-Is parasitemia significantly higher in birds sampled in urban or rural environments?
-These are the main variables I will be looking at for this project, but I may analyze others as the project continues.
-Not all of the variables were collected for every Ibis sampled, and the data still needs cleaning to accurately detail the sample size for each analysis.
-
-Is there a difference in HaeParasit between Ibis of different ages?
-- Dependent variable (outcome): HaeParasit
-- Independent variable (predictor): Age
-
- Is there a difference in HaeParasit between Ibis of different sex?
-- Dependent variable (outcome): HaeParasit
-- Independent variable (predictor): Sex
-
-Is there a different in HaeParasit in Ibis sampled at different sites?
- - Dependent variable (outcome): HaeParasit
- - Independent variable (predictor): Site
-
- Is there a difference in HaeParasit between Ibis sampled in different habitat types?
- - Dependent variable (outcome): HaeParasit
- - Independent variable (predictor): HabType
-
- Is there a difference in HaeParasit between Ibis sampled at different times?
- - Dependent variable (outcome): HaeParasit
- - Independent variables (predictors): Date, Season
-
-## Expected Analyses
-I think that I will have to conduct multiple ANOVA's during the analysis process.
-Linear and logistic regression models apply here too for some variables.I would like to explore the possibility of using the location data for each sample to conduct a geographical analysis with a map as the output.
-I will be continuing to contact those involved with the study as well to see what analyses they want to be conducted and what questions they want answered.
+The data was collected by various undergraduate and graduate students of Dr.
+Sonia Hernandez and Dr.
+Michael Yabsley at the University of Georgia.
+Wild American White Ibis were captured during multiple sampling seasons between 2010-2017 in various urban and natural habitat sites from southern Florida.
+Data was recorded in excel sheets that included information about the sampling sites (habitat type, location, etc.) and various characteristics about the captured Ibis (weight, sex, age, etc.).
+Ibis were bled and lab diagnostics were performed at the Southeastern Cooperative Wildlife Disease Study (SCWDS) at UGA in order to determine the presence of *Haemoproteus* parasitemia, and if possible, amount of parasitemia in the blood.
 
 ## Data Cleaning
-The data has already been imported, so now we move onto the cleaning stage.
-The following code will clean, process, and save the raw Ibis blood datasets from 2010-2014 and 2015-2017.
-This section is contained in the "ProcessingScriptIbisBloodData.R" file.
-```{r}
-##############################Cleaning the rdIbisBlood10_14 data###########################
-IbisBlood10_14 <- rename(rdIbisBlood10_14, ID = X.) #renames the ID column
 
-#Extracts just the columns of interest
-IbisBlood10_14 <- select(IbisBlood10_14, ID, Site.ID, Date, PCR.sex, Age..adult.juvenile., 
-                         Total.RBC, Haemoproteus.parasitemia.)
+The data was organized in 2 excel sheets; one containing all the lab diagnostic data concerning the sampled Ibis and the other containing all the descriptive, demographic, and geographic data about the sampled Ibis.
+Both excel files were combined while removing uninformative and duplicate variables in both.
+Variables of importance were standardized once both datasheets were combined.
+If sampling season was not recorded but sampling date was for an Ibis, season was determined from the date and added.
+I dichotomized the recorded amount of *Haemoproteus* parasitemia to either "1" (presence) or "0" (absence) by whether or not the Ibis had a parasitemia value of above zero.
+Locations that did not have presence of *Haemoproteus* and had less than 5 observations with little demographic information were removed from analysis.
 
-#Renames the columns of interest
-IbisBlood10_14 <- rename(IbisBlood10_14, Site = Site.ID, Sex = PCR.sex, Age = Age..adult.juvenile.,
-                         TotalRBC = Total.RBC, HaeParasit = Haemoproteus.parasitemia.)
+## Data Analysis
 
-#Change "adult" and "juvenile" in the Age column to "A" and "J"
-IbisBlood10_14$Age[IbisBlood10_14$Age == "adult"] <- "A"
-IbisBlood10_14$Age[IbisBlood10_14$Age == "Adult"] <- "A"
-IbisBlood10_14$Age[IbisBlood10_14$Age == "juvenile"] <- "J"
-IbisBlood10_14$Age[IbisBlood10_14$Age == "SA"] <- "A"
-#Now the Ibises that had age recorded are referred to as "A" for adult or "J" for juvenile
+Univariate relationships between the presence of *Haemoproteus* parasitemia and each predictor were assessed using logistic regression treating *Haemoproteus* parasitemia presence as a categorical outcome.
+Multivariate classification was also performed, determination of variables to include was performed by sequential forward floating selection.
+In addition, a Least Absolute Shrinkage and Selection Operator (LASSO) model, a decision tree, and a random forest model were all conducted with *Haemoproteus* parasitemia presence as the outcome variable.
+The goal of all models was maximization of accuracy and roc_auc on test data using a 10 fold cross-validation (80/20 split, 5 times repeated).
 
-#Changing the DIV/0 values in HaeParasit to NA and change the column to numerical values
-IbisBlood10_14$HaeParasit <- as.numeric(IbisBlood10_14$HaeParasit)  # Divide-by-zero errors are now NA
+## Study Population
 
-#Reformatting the date values in the Date column
-IbisBlood10_14$Date <- excel_numeric_to_date(IbisBlood10_14$Date)
-#After this step, the values that just state the year are labelled with "1905", may need to go back or exclude from analysis
+During all of the sampling seasons, 668 Ibis were sampled.
+Among those sampled, 355 Ibis had a positive *Haemoproteus* parasitemia value.
+20 Ibis were removed from the analysis because they did not have a positive *Haemoproteus* parasitemia value and the site from which they were captured had less than or equal to 5 sampled Ibis and had no presence of infected Ibis.
 
-#Abbreviating the site names to make analysis easier
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Royal Palms"] <- "RP"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Everglades"] <- "E"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Dreher Park"] <- "DRP"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Indian Creek"] <- "ICP"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Indian Creek "] <- "ICP"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Indian Creek Park"] <- "ICP"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Juno Beach"] <- "JB"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Lion Country Safari"] <- "LCS"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Lake Worth"] <- "LW"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Prosperity Oaks"] <- "PO"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Promenade Plaza"] <- "PP"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "San Mateo"] <- "SM"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "Solid Waste Authority"] <- "SWA"
-IbisBlood10_14$Site[IbisBlood10_14$Site == "West Palm Beach Zoo"] <- "WPBZ"
-
-#Reformatting the mislabeled dates. These are set to Jan 1st of the year they were
-#recorded on the data sheet, these entries only had the year stated, not month and day.
-IbisBlood10_14$Date[IbisBlood10_14$Date == "1905-07-02"] <- "2010-01-01"
-IbisBlood10_14$Date[IbisBlood10_14$Date == "1905-07-03"] <- "2011-01-01"
-IbisBlood10_14$Date[IbisBlood10_14$Date == "1905-07-04"] <- "2012-01-01"
-
-#The bottom of the dataset has a couple rows of just missing data for all variables,
-#This code filters the dataset so that if any rows are present that has at least
-#one variable be NOT NA, R keeps it
-IbisBlood10_14 <- filter(IbisBlood10_14, if_any(everything(), ~ !is.na(.)))
-
-#This completes the cleaning phase for IbisBlood10_14
-#------------------------------------------------------------------------------------#
-
-############################Cleaning the rdIbisBlood15_17 Data#########################
-
-IbisBlood15_17<- rename(rdIbisBlood15_17, ID= X.) #renames the ID column
-
-####Extracting only the columns relevant for analysis
-IbisBlood15_17 <- select(IbisBlood15_17,ID,Site.Name,Date,PCR.sex,Age,Season,Habitat.type,
-                         Total.RBC,X..Haemoproteus,Haemoproteus.parasitemia.,F1,F2,F3,
-                         Avg.RBC.FOV,Ibis..,Total.mass..g.,Mass.bag..g.,Mass.bird..g.,Body.condition.score..1.5.,
-                         Ectoparasite.score..1.5.,Culmen.length..mm.,Wing.chord.length..mm.,
-                         Tarsus.length..mm.,Tarsus.width..mm.)
-
-####Renaming the columns 
-IbisBlood15_17 <- rename(IbisBlood15_17,Site = Site.Name,Sex = PCR.sex,HabType = Habitat.type,TotalRBC = Total.RBC,NumHae = X..Haemoproteus,
-                         HaeParasit = Haemoproteus.parasitemia.,AvgRBC = Avg.RBC.FOV,
-                         IbisNum = Ibis..,TotalMassG = Total.mass..g.,BagMassG = Mass.bag..g.,
-                         BirdMassG = Mass.bird..g.,BodyCondScore = Body.condition.score..1.5.,
-                         EctoParasitScore = Ectoparasite.score..1.5.,CulmenLmm = Culmen.length..mm.,
-                         WingChordLmm = Wing.chord.length..mm.,TarsusLmm = Tarsus.length..mm.,
-                         TarsusWmm = Tarsus.width..mm.)
-
-####Site column
-#Ibis that have sites matched with the 10_14 dataset will have the same labels
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Juno Beach"] <- "JB"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Indian Creek Park"] <- "ICP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Dubois park"] <- "DUP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Dreher Park"] <- "DRP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Lion Country Safari"] <- "LCS"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Dubois Park"] <- "DUP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "J.W. Corbett Wildlife Management Area"] <- "JWC"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Gaines Park"] <- "GP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Solid Waste Authority"] <- "SWA"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Fisheating creek"] <- "FC"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Kitching Creek"] <- "KC"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Loxahatchee NE"] <- "LOXNE"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "DuBois Park"] <- "DUP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "TetraTech"] <- "TT"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Green Cay"] <- "GC"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Dreher park"] <- "DRP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "CROW Wildlife Rehabilitation Center"] <- "CROW"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Lake Worth"] <- "LW"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Royal Palm"] <- "RP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "JW Corbett"] <- "JWC"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Busch Wildlife Sanctuary"] <- "BWS"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Loxahatchee Slough"] <- "LOXS"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Indian Creek"] <- "ICP"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Cat House (Bonnie's, Richard Lane)"] <- "CAT"
-IbisBlood15_17$Site[IbisBlood15_17$Site == "Palm Beach Zoo"] <- "WPBZ"
-
-####Date column
-#Changing the date column to numeric values to allow date extraction
-IbisBlood15_17$Date <- as.numeric(IbisBlood15_17$Date) 
-##Reformatting the date values in the Date column
-IbisBlood15_17$Date <- excel_numeric_to_date(IbisBlood15_17$Date)
-
-####Sex column
-IbisBlood15_17$Sex[IbisBlood15_17$Sex == "not bled"] <- NA
-#Two values have missing data because the birds were not bled to determine sex
-IbisBlood15_17$Sex[IbisBlood15_17$Sex == "M?"] <- NA
-IbisBlood15_17$Sex[IbisBlood15_17$Sex == "F?"] <- NA
-IbisBlood15_17$Sex[IbisBlood15_17$Sex == "F "] <- "F"
-#Two values have question marks for unsure sex determination, removing those values
-
-####Age column
-#When an ibis reaches 4 years old, their plumage becomes all white. Any Ibis
-#that was sampled with brown plumage is considered a juvenile. Since Haemoproteus is
-#a chronic disease that mostly likely accumulates over time, long-lived birds should
-#theoretically have more parasitemia. Could not be true though and may come back to
-#look at the individual years in the juvenile birds
-IbisBlood15_17$Age[IbisBlood15_17$Age == "Juvenile"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "Adult"] <- "A"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "2nd/3rd year"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "3rd year"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "1st year"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "adult"] <- "A"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "2nd year"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "NA"] <- NA
-IbisBlood15_17$Age[IbisBlood15_17$Age == "2nd Year"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "Late third year"] <- "J"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "ADULT"] <- "A"
-IbisBlood15_17$Age[IbisBlood15_17$Age == "Adult bc"] <- "A" #Not sure what bc stands for
-
-#Season and HabType are fine the way they are
-
-####TotalRBC
-IbisBlood15_17$TotalRBC <- as.numeric(IbisBlood15_17$TotalRBC)
-#DIV/0 errors are now NA in this column
-
-#NumHae column is fine
-
-####HaeParasit
-IbisBlood15_17$HaeParasit <- as.numeric(IbisBlood15_17$HaeParasit)
-
-####-999 values
-IbisBlood15_17[IbisBlood15_17 == -999] <- NA
-
-####Removing the last couple empty rows
-IbisBlood15_17 <- IbisBlood15_17[-c(401, 402, 403, 404, 405, 406),]
-
-#This completes the cleaning phase for IbisBlood15_17
-#------------------------------------------------------------------------------------#
-
-###############################Saving the Cleaned Datasets###########################
-
-#Saving the Ibis Blood 10-14 dataset as an RDS in the processed data folder
-save_data_location <- here::here("processed_data", "processedIbisBlood10_14.rds")
-saveRDS(IbisBlood10_14, file = save_data_location)
-
-#Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
-save_data_location2 <- here::here("processed_data", "processedIbisBlood15_17.rds")
-saveRDS(IbisBlood15_17, file = save_data_location2)
-
-```
-
-This next section will clean, process, and save the raw Ibis Field dataset. This section is contained in the "ProcessingScriptIbisFieldData.R" file.
-
-```{r}
-######################### #Cleaning the rdIbisField Data###############################
-
-IbisField <- rename(rdIbisField, ID = ID..NSFSITE...) #renames the ID column
-
-#Selecting the columns of interest to this analysis from the datasheet
-IbisField <- select(IbisField,ID,Site.Name,Latitude,Longitude,Date,PCR.Sex,Age,Season,
-                    Habitat.type,Ibis.number,Mass.bird..g.,Body.condition.score..1.5.,
-                    Ectoparasite.score..1.5.,Culmen.length..mm.,Wing.chord.length..mm.,
-                    Tarsus.length..mm.,Tarsus.width..mm.,Habituation.score.of.flock..1.5.)
-
-#Renaming the columns of interest, also matching the other cleaned datasets
-IbisField <- rename(IbisField,Site = Site.Name,Lat = Latitude,Long = Longitude,Sex = PCR.Sex,
-                    HabType = Habitat.type,IbisNum = Ibis.number,BirdMassG = Mass.bird..g.,
-                    BodyCondScore = Body.condition.score..1.5.,EctoParasitScore = Ectoparasite.score..1.5.,
-                    CulmenLmm = Culmen.length..mm.,WingChordLmm = Wing.chord.length..mm.,
-                    TarsusLmm = Tarsus.length..mm.,TarsusWmm = Tarsus.width..mm.,HabFlockScore = Habituation.score.of.flock..1.5.)
-
-#This dataset has "-999" values which mean they are intentionally left blank,
-#Birds with an ID number of "-999" represent sites that didn't catch any birds,
-#but had weather data collected. This code deletes those observations.
-IbisField <- IbisField[!(IbisField$ID == "-999"),]
-
-####Site
-#These site location abbreviations are the same across datasets
-IbisField$Site[IbisField$Site == "Busch Wildlife Sanctuary"] <- "BWS"
-IbisField$Site[IbisField$Site == "Cat House (Bonnie's, Richard Lane)"] <- "CAT"
-IbisField$Site[IbisField$Site == "Dubois Park"] <- "DUP"
-IbisField$Site[IbisField$Site == "DuBois Park"] <- "DUP"
-IbisField$Site[IbisField$Site == "Dreher Park"] <- "DRP"
-IbisField$Site[IbisField$Site == "Fisheating Creek"] <- "FC"
-IbisField$Site[IbisField$Site == "Green Cay"] <- "GC"
-IbisField$Site[IbisField$Site == "Gaines Park"] <- "GP"
-IbisField$Site[IbisField$Site == "Indian Creek"] <- "ICP"
-IbisField$Site[IbisField$Site == "Juno Beach"] <- "JB"
-IbisField$Site[IbisField$Site == "Juno Beach "] <- "JB"
-IbisField$Site[IbisField$Site == "J.W. Corbett Wildlife Management Area"] <- "JWC"
-IbisField$Site[IbisField$Site == "Kitching Creek"] <- "KC"
-IbisField$Site[IbisField$Site == "Lion Country Safari"] <- "LCS"
-IbisField$Site[IbisField$Site == "Lake Worth"] <- "LW"
-IbisField$Site[IbisField$Site == "Loxahatchee Slough"] <- "LOXS"
-IbisField$Site[IbisField$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
-IbisField$Site[IbisField$Site == "Loxahatchee/LILA Cell B2"] <- "LOXB2"
-IbisField$Site[IbisField$Site == "Loxahatchee NE "] <- "LOXNE"
-IbisField$Site[IbisField$Site == "Palm Beach Zoo"] <- "WPBZ"
-IbisField$Site[IbisField$Site == "Royal Palm"] <- "RP"
-IbisField$Site[IbisField$Site == "CROW Wildlife Rehabilitation Center"] <- "CROW"
-IbisField$Site[IbisField$Site == "Solid Waste Authority"] <- "SWA"
-IbisField$Site[IbisField$Site == "TetraTech"] <- "TT"
-
-####Sex Column
-#This replaces the -999 and FAIL values with NA. The IbisBlood datasheets carry more sex data for the birds
-IbisField$Sex[IbisField$Sex == "FAIL"] <- NA
-
-####Age Column
-IbisField$Age[IbisField$Age == "Adult"] <- "A"
-IbisField$Age[IbisField$Age == "3rd year"] <- "J"
-IbisField$Age[IbisField$Age == "2nd year"] <- "J"
-IbisField$Age[IbisField$Age == "1st year"] <- "J"
-IbisField$Age[IbisField$Age == "NA"] <- NA
-IbisField$Age[IbisField$Age == "Adult "] <- "A"
-IbisField$Age[IbisField$Age == "2nd year "] <- "J"
-IbisField$Age[IbisField$Age == "Juvenile"] <- "J"
-IbisField$Age[IbisField$Age == "2nd Year"] <- "J"
-
-#Replacing all -999 values in the dataset with NA
-IbisField[IbisField == -999] <- NA
-
-#Replacing NA text strings with regular NA values in multiple columns
-IbisField$BodyCondScore[IbisField$BodyCondScore == "NA"] <- NA
-IbisField$HabFlockScore[IbisField$HabFlockScore == "NA"] <- NA
-IbisField$CulmenLmm[IbisField$CulmenLmm == "NA"] <- NA
-IbisField$WingChordLmm[IbisField$WingChordLmm == "NA"] <- NA
-IbisField$TarsusLmm[IbisField$TarsusLmm == "NA"] <- NA
-IbisField$TarsusWmm[IbisField$TarsusWmm == "NA"] <- NA
-
-#--------------------------------------------------------------------------------------#
-
-###############################Saving the Cleaned Dataset###############################
-
-save_data_location3 <- here::here("processed_data", "processedIbisFielddata.rds")
-saveRDS(IbisField, file = save_data_location3)
-```
-
-This cleaning code section will clean, process, and save the Ibis urbanization dataset. This section is contained in the "ProcessingScriptIbisUrban.R" file.
-
-```{r}
-
-##############################Cleaning the rdUrbanIbis data###########################
-#Selects only the columns of interest in the dataset
-IbisUrban <- select(rdUrbanIbis, Name:Serotype)
-
-#Renames the columns of interest
-IbisUrban <- rename(IbisUrban,ID = Name,Date = Collection.Date,HgPPM = mg.kg.Hg..PPM.,
-                    Site = Site.Name,UrbanPercent = Site...Urbanized)
-
-#Change the site names to their abbreviations
-IbisUrban$Site[IbisUrban$Site == "Juno Beach"] <- "JB"
-IbisUrban$Site[IbisUrban$Site == "Indian Creek"] <- "ICP"
-IbisUrban$Site[IbisUrban$Site == "Dubois Park"] <- "DUP"
-IbisUrban$Site[IbisUrban$Site == "Dreher Park"] <- "DRP"
-IbisUrban$Site[IbisUrban$Site == "Lion Country Safari"] <- "LCS"
-IbisUrban$Site[IbisUrban$Site == "Loxahatchee Wildlife Refuge"] <- "LOXWR"
-IbisUrban$Site[IbisUrban$Site == "Solid Waste Authority "] <- "SWA"
-IbisUrban$Site[IbisUrban$Site == "Gaines Park"] <- "GP"
-IbisUrban$Site[IbisUrban$Site == "Kitching Creek"] <- "KC"
-IbisUrban$Site[IbisUrban$Site == "Kitching Creek "] <- "KC"
-IbisUrban$Site[IbisUrban$Site == "Loxahatchee NE"] <- "LOXNE"
-IbisUrban$Site[IbisUrban$Site == "TetraTech"] <- "TT"
-IbisUrban$Site[IbisUrban$Site == "Loxahatchee NE "] <- "LOXNE"
-IbisUrban$Site[IbisUrban$Site == "Green Cay"] <- "GC"
-IbisUrban$Site[IbisUrban$Site == "J.W. Corbett Wildlife Management Area"] <- "JWC"
-IbisUrban$Site[IbisUrban$Site == "Loxahatchee Slough"] <- "LOXS"
-IbisUrban$Site[IbisUrban$Site == "Lake Worth"] <- "LW"
-
-#--------------------------------------------------------------------------------------#
-
-################################Saving the UrbanIbis Data#############################
-#Saving the Ibis Blood 15-17 dataset as an RDS in the processed data folder
-save_data_location3 <- here::here("processed_data", "processedIbisUrban.rds")
-saveRDS(IbisUrban, file = save_data_location3)
-
-#--------------------------------------------------------------------------------------#
-```
-
-This last cleaning section combines the data sheets into one to use for analysis.
-The urbanization gradient has not been created as of yet, but a meeting this month with those involved with the project will help shed some light on the exact specifics for analysis.
-This code is located in the "ExploratoryAnalysisScript.R" file.
-
-```{r}
-###############################Completing Combination Data Frames#########################
-#Outer joining IbisBlood10_14 and IbisBlood15_17, keeping all values of the 10_14 dataset
-IbisBlood10_17 <- merge(IbisBlood10_14, IbisBlood15_17, by = c("ID", "Site", "Date", "Sex", "Age",
-                                                              "TotalRBC", "HaeParasit"), all=TRUE)
-#There is no overlap in the ID numbers between the two, so no recapture is assumed.
-
-#This is a left join, this data frame below only contains birds that were in the IbisBlood10_17
-#dataset (all with parasitemia data), with additional information added by the IbisField data. Ibis
-#from the Field data that weren't tested for parasitemia or weren't included on Blood dataset
-#are not included in this dataframe
-
-#IbisBlood10_17(with Field)_(Parasitemia data)
-IbisC <- left_join(IbisBlood10_17, IbisField, by = c("ID", "Site", "Date", "Sex", "Age", "Season", "HabType",
-                                                         "IbisNum", "BirdMassG", "BodyCondScore", "EctoParasitScore",
-                                                         "CulmenLmm", "WingChordLmm", "TarsusLmm", "TarsusWmm"))
-
-#This section classifies the habitat type for the sites without a habitat type.
-IbisC <- IbisC %>%
-        mutate(HabType = ifelse(Site == "ICP", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "JB", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "RP", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "LCS", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "LW", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "SWA", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "E", "Natural", HabType)) %>%
-        mutate(HabType = ifelse(Site == "WPBZ", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "PO", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "PP", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "DRP", "Urban", HabType)) %>%
-        mutate(HabType = ifelse(Site == "SM", "Urban", HabType))
-
-##This section is adding in the seasons for observations that have collection dates but no season.
-
-#Creating intervals for adding seasons based on the date column
-int1 <- interval(ymd("2012-12-01"), ymd("2012-12-31"))
-int2 <- interval(ymd("2013-06-01"), ymd("2013-08-31"))
-int3 <- interval(ymd("2013-12-01"), ymd("2013-12-31"))
-int4 <- interval(ymd("2014-02-01"), ymd("2014-03-31"))
-
-#Adding seasons to the ibis without them stated based on the date the sameple was collected
-IbisC <- IbisC %>%
-         mutate(Season = ifelse(Date %within% int1, "Fall 2012", Season)) %>%
-         mutate(Season = ifelse(Date %within% int2, "Summer 2013", Season)) %>%
-         mutate(Season = ifelse(Date %within% int3, "Fall 2013", Season)) %>%
-         mutate(Season = ifelse(Date %within% int4, "Spring 2014", Season))
-
-#Adding a column for HaeParasit Presence/Absence data
-IbisC <- IbisC %>%
-    mutate(HaeParasitPA = as.integer(HaeParasit > 0 & !is.na(HaeParasit)))
-
-#Adding a column for transformed HaeParasit data on a logarithm base 10 scale,
-#then removing the -Inf values due to there being no parasitemia in those birds. This column
-#then only contains logarithm transformed parasitemia data for birds with a positive value
-#of parasitemia.
-IbisC <- IbisC %>%
-    mutate(HaeParasitLog10 = log10(HaeParasit)) %>%
-    mutate(HaeParasitLog10 = ifelse(HaeParasitLog10 == "-Inf", NA, HaeParasitLog10))
-
-#Changing the BodyCondScore to an ordinal scale
-IbisC <- IbisC %>%
-      mutate(BodyCondScore = recode_factor(BodyCondScore, '1' = 'Emaciated',
-                                           '2' = 'Underweight', '3' = 'Ideal',
-                                           '4' = 'Overweight', '5' = 'Obese',
-                                           .ordered = TRUE))
-
-#Changing the class of some variables for analysis. From character to numeric.
-C2N <- c(16,17,18,21,22,23,24)
-
-IbisC[ , C2N] <- apply(IbisC[ , C2N], 2,
-                       function(x) as.numeric(as.character(x)))
-
-```
-
-This concludes the data cleaning section, we are now ready for the exploratory analysis!
-
-## Exploratory analysis
+## Exploratory Analysis
 

### 2021-11-11 analysis_scripts/ExploratoryAnalysisScript.R

@@ -19,12 +19,10 @@ library(lubridate) #work with date/time data
 data_location1 <- here::here("processed_data","processedIbisBlood10_14.rds")
 data_location2 <- here::here("processed_data","processedIbisBlood15_17.rds")
 data_location3 <- here::here("processed_data","processedIbisFielddata.rds")
-data_location4 <- here::here("processed_data","processedIbisUrban.rds")
 
 IbisBlood10_14 <- readRDS(data_location1)
 IbisBlood15_17 <- readRDS(data_location2)
 IbisField <- readRDS(data_location3)
-IbisUrban <- readRDS(data_location4)
 
 #--------------------------------------------------------------------------------------#
 
@@ -97,11 +95,17 @@ IbisC <- IbisC %>%
                                            '4' = 'Overweight', '5' = 'Obese',
                                            .ordered = TRUE))
       
-#Changing the class of some variables for analysis. From character to numeric.
+#Changing the class of some variables for graphing. From character to numeric.
 C2N <- c(16,17,18,21,22,23,24)
 
 IbisC[ , C2N] <- apply(IbisC[ , C2N], 2,
                        function(x) as.numeric(as.character(x)))
+
+#Deleting sites that hinder the model, these sites have no presence of Haemoproteus #ERROR: factor Site has new levels CAT, E, SM, WPBZ
+IbisC <- subset(IbisC, Site!="CAT" &
+                     Site!="E" &
+                     Site!="SM" &
+                     Site!="WPBZ" )
 
 ##############################Saving the Complete Dataset#########################
 
@@ -110,6 +114,9 @@ save_data_location <- here::here("processed_data", "CompleteIbisDataset.rds")
 saveRDS(IbisC, file = save_data_location)
 
 ###############################Exploratory Tables#################################
+IbisC %>%
+  count(HaeParasitPA, wt = HaeParasitPA)
+
 IbisC %>%
 count(Sex, wt = HaeParasitPA)
 
@@ -129,7 +136,7 @@ IbisC %>%
   count(BodyCondScore, Age, wt = HaeParasitPA)
 
 ################################Exploratory Graphs#####################################
-########Plot 1
+########Plot 1: Scatter Plot of Haemoproteus parasitemia, sample date and habitat type
 ExpFigure1 <-ggplot(IbisC) +
   geom_point(aes(x = Date, y = HaeParasitLog10, color = HabType)) +
   ggtitle("Haemoproteus parasitemia from sampled infected Ibises from 2010-2017 by habitat type") +

### 2021-11-11 analysis_scripts/StatisticalAnalysisModels.R

@@ -10,7 +10,9 @@ library(tidyverse) #all required data manipulation packages
 library(lubridate) #work with date/time data
 library(tidymodels) #to run the statistical analyses
 library(ranger) #implementation of random forests
-
+library(vip) #For plotting decision tree model
+library(rpart) #For plotting machine learning models
+library(rpart.plot) #For plotting machine learning models
 set.seed(144)
 #--------------------------------------------------------------------------------------#
 
@@ -40,13 +42,7 @@ IbisPAdf$HabType <- as.factor(IbisPAdf$HabType)
 #Changing date to numeric for the model
 IbisPAdf$Date <- as.numeric(IbisPAdf$Date)
 
-#Deleting sites that hinder the model, these sites have no presence of Haemoproteus #ERROR: factor Site has new levels CAT, E, SM, WPBZ
-IbisPAdf <- subset(IbisPAdf, Site!="CAT" &
-                     Site!="E" &
-                     Site!="SM" &
-                     Site!="WPBZ" )
-
-#####################################Cross Validation and Model Preparation
+#####################################Cross Validation Preperation#####################################
 ###Cross-Validation Setup
 #Splitting the data in an 80/20 proportion
 data_split_log <- initial_split(IbisPAdf, 0.8)
@@ -57,22 +53,54 @@ PATestDF <- testing(data_split_log) #Creating the testing dataset
 
 #Resampling the data with 10-fold cross validation. Splits the training
 #data into 10 samples to train the model
-PAFolds <- vfold_cv(PATrainDF, v = 10)
+PAFolds <- vfold_cv(PATrainDF, v = 10, repeats = 5)
 
-#Binomial GLM model setup
-log_mod1 <- logistic_reg() %>% set_engine("glm") %>% set_mode("classification") 
+#####################################Null Model Setup################################################
 
-########################################Model Evaluation
+#sets the basic linear model for training data
[…]

### 2021-11-11 processing_script/ProcessingScript.R

@@ -16,16 +16,12 @@ library(janitor) #to help clean up dates and tidy data
 data_locationIbisB <- here::here("raw_data","Ibis_Blood_ParasiteData_2017.xlsx")
 #Setting the path to the Ibis field data spreadsheet
 data_locationIbisF <- here::here("raw_data","IbisFieldDataOct19_2017.xlsx")
-#Setting the path to the first Ibis supplemental excel sheet (Urban Gradients)
-data_locationIbisUr <- here::here("raw_data","100IbisID_UrbanGradient.xlsx")
 #Loads the raw Ibis blood data excel sheet from 2015-2017
 rdIbisBlood15_17 <- read.xlsx(data_locationIbisB, sheetName = "2015-2017")
 #Loads the raw Ibis blood data excel sheet from 2010-2014
 rdIbisBlood10_14 <- read.xlsx(data_locationIbisB, sheetName = "2010-2014")
 #Loads the raw Ibis field data excel sheet
 rdIbisField <- read.xlsx(data_locationIbisF, sheetName = "All capture data")
-#Loads the raw Ibis Urban Gradient supplemental data sheet
-rdUrbanIbis <- read.xlsx(data_locationIbisUr, sheetName = "Samples by Date")
 #--------------------------------------------------------------------------------------#
 
 ##############################Cleaning the rdIbisBlood10_14 data###########################
@@ -265,35 +261,6 @@ IbisField$TarsusWmm[IbisField$TarsusWmm == "NA"] <- NA
 
 #--------------------------------------------------------------------------------------#
 
-##############################Cleaning the rdUrbanIbis data###########################
-#Selects only the columns of interest in the dataset
-IbisUrban <- select(rdUrbanIbis, Name:Serotype)
-
-#Renames the columns of interest
-IbisUrban <- rename(IbisUrban,ID = Name,Date = Collection.Date,HgPPM = mg.kg.Hg..PPM.,
-                    Site = Site.Name,UrbanPercent = Site...Urbanized)
-
-#Change the site names to their abbreviations
-IbisUrban$Site[IbisUrban$Site == "Juno Beach"] <- "JB"
-IbisUrban$Site[IbisUrban$Site == "Indian Creek"] <- "ICP"
[…]

### 2021-11-13 products/manuscript/Manuscript.Rmd
Update Manuscript Part 4
Updated manuscript with new tables and figures.

@@
 library(lubridate) #To extract times from raw data
+library(tidymodels) #to run the statistical analyses
+library(ranger) #implementation of random forests
+library(vip) #For plotting decision tree model
+library(rpart) #For plotting machine learning models
+library(rpart.plot) #For plotting machine learning models
+set.seed(144)
+#Loading in the complete Ibis Dataset
+data_location1 <- here::here("processed_data","CompleteIbisDataset.rds")
+IbisC <- readRDS(data_location1)
+#Loading required materials - Figures
+#Exploratory Figures
+figure_Exp_1 <- here::here("results", "Figures", "ExpFigure1.png")
+figure_Exp_4 <- here::here("results", "Figures", "ExpFigure4.png")
+figure_Exp_6 <- here::here("results", "Figures", "ExpFigure6.png")
+#Machine Learning Model Diagnostic Figures
+figure_ML_1 <- here::here("results", "Figures", "ImpPlotDT.jpeg")
+figure_ML_2 <- here::here("results", "Figures", "DTTree.jpeg")
+figure_ML_3 <- here::here("results", "Figures", "LassoFig.jpeg")
+figure_ML_4 <- here::here("results", "Figures", "ImpPlotRF.jpeg")
+figure_ML_5 <- here::here("results", "Figures", "ROC_Test_RF.jpeg")
+#Loading and preparing the Exploratory Analysis Tables
+IbisCount1 <- IbisC %>% count(Sex, Age, wt = HaeParasitPA)
+IbisCount2 <- IbisC %>% count(Season, wt = HaeParasitPA)
+IbisCount3 <- IbisC %>% count(Site, wt = HaeParasitPA)
+#Loading Model Results
+data_locationPA1 <- here::here("results", "Tables", "Model1MetricsPA.Rda")
+model1 <- readRDS(data_locationPA1)
+data_locationPA2 <- here::here("results", "Tables", "Model2MetricsPA.Rda")
+model2 <- readRDS(data_locationPA2)
+data_locationPA3 <- here::here("results", "Tables", "Model3MetricsPA.Rda")
+model3 <- readRDS(data_locationPA3)
+data_locationPA4 <- here::here("results", "Tables", "Model4MetricsPA.Rda")
+model4 <- readRDS(data_locationPA4)
+data_locationPA5 <- here::here("results", "Tables", "Model5MetricsPA.Rda")
+model5 <- readRDS(data_locationPA5)
+data_locationPA6 <- here::here("results", "Tables", "Model6MetricsPA.Rda")
+model6 <- readRDS(data_locationPA6)
+data_locationPA7 <- here::here("results", "Tables", "Model7MetricsPA.Rda")
+model7 <- readRDS(data_locationPA7)
+data_locationPA8 <- here::here("results", "Tables", "Model8MetricsPA.Rda")
+model8 <- readRDS(data_locationPA8)
+data_locationPA9 <- here::here("results", "Tables", "Model9MetricsPA.Rda")
+model9 <- readRDS(data_locationPA9)
+#Loading required materials - Tables
+mod_loc_1 <- here::here("results", "Models", "Model1PA.Rda")
+mod_loc_2 <- here::here("results", "Models", "Model2PA.Rda")
+mod_loc_3 <- here::here("results", "Models", "Model3PA.Rda")
+mod_loc_4 <- here::here("results", "Models", "Model4PA.Rda")
+mod_loc_5 <- here::here("results", "Models", "Model5PA.Rda")
+mod_loc_6 <- here::here("results", "Models", "Model6PA.Rda")
+mod_loc_7 <- here::here("results", "Models", "Model7PA.Rda")
+mod_loc_8 <- here::here("results", "Models", "Model8PA.Rda")
+mod_loc_9 <- here::here("results", "Models", "Model9PA.Rda")
+
 ```
 
-# Introduction
+## Introduction
 
@@
 The parasite the data focuses on is the protozoan *Haemoproteus*, which causes a chronic blood infection in White Ibises and many other birds.
-*Haemoproteus* is a vector-borne disease, it is transmitted by *Culicoides* biting midges when they feed on the birds and causes long-lasting parasitemia.
-Adult healthy birds tolerate haemosporidian infections well, but young birds and birds outside their normal range that are naive to the pathogen are more likely to develop clinical disease.
-Haemosporidian infections usually cause long-lasting chronic infections in natural hosts, and can cause reduced reproductive success, host fitness, and increased stress and disease susceptibility.
+*Haemoproteus* is a vector-borne disease, it is transmitted by *Culicoides* biting midges when they feed on the birds and causes long-lasting parasitemia [@Coker2017].
+Adult healthy birds tolerate haemosporidian infections well, but young birds and birds outside their normal range that are naive to the pathogen are more likely to develop clinical disease [@Coker2017].
+Haemosporidian infections usually cause long-lasting chronic infections in natural hosts, and can cause reduced reproductive success, host fitness, and increased stress and disease susceptibility [@Coker2017].
 
@@
 
-# Methods
+## Methods
 
-## Data Collection
+### Data Collection
 
@@
 
-## Data Cleaning
+### Data Cleaning
 
@@
 
-## Data Analysis
+### Data Analysis
 
@@
 
-## Study Population
+### Study Population
 
@@
 20 Ibis were removed from the analysis because they did not have a positive *Haemoproteus* parasitemia value and the site from which they were captured had less than or equal to 5 sampled Ibis and had no presence of infected Ibis.
+Ibis were either classified as having parasitemia (1 - presence) or not having parasitemia (0 - absence).
 
-## Exploratory Analysis
+### Exploratory Analysis
 
-### Tables
+Here we looked at various variate and covariate relationships between the predictor variables and the main outcome variable of presence of *Haemoproteus* in the sampled Ibis.
+The following tables and figures show those relationships prior to the statistical analysis.
 
-To start the exploratory analysis, we're going to create some quick tables to get an idea of the metrics for some of the columns of interest.
+#### Tables
+```{r}
+IbisCount1
+```
+Overall, looking at the distribution of Ibis Sex and Age, the are more females than males and adults than juveniles.
+Female adults have the highest number with 109.
+Another clear observation from this table is that there are a large number of these missing variables in the dataset.
+This will most likely have to be completed with and without imputation depending on the model used.
 
 ```{r}
-IbisC %>%
-  count(Sex, wt = HaeParasitPA)
+IbisCount2
 ```
-
-Looking at presence/absence data for haemoproteus parasitemia, there were about double the number of female ibises with active infection compared to the males.
-78 of the ibises sampled were negative for haemoproteus infection.
+This table shows the number of Ibis sampled during each season. The majority of the Ibis were sampled in the later years (2015-2017) than in the earlier years (2010-2014).
 
 ```{r}
-IbisC %>%
-  count(Age, wt = HaeParasitPA)
+IbisCount3
 ```
+The majority of the Ibis were sampled in SWA, DRP, and ICP sites. Most of the Ibis were sampled in urban sites compared to natural sites.
 
-Here we can see that there was about 2.7 times more adults sampled with active infection than juvenile ibises.
-59 of the sampled ibises with active infection either couldn't be aged or age was not recorded.
-
-```{r}
-IbisC %>%
-  count(Age, Sex, wt = HaeParasitPA)
-```
-
-Combining age and sex, we can see that adult females had the highest number of active infections. Categories with missing numbers are prevalent.
-
-```{r}
-IbisC %>%
-  count(Site, wt = HaeParasitPA)
-```
-Looking at site, SWA, ICP, and DRP were the sites with the highest number of sampled ibises with active infection.
-
-```{r}
-IbisC %>%
-  count(Season, wt = HaeParasitPA)
-```
-Looking at season sampled, Spring 2016 had the highest number of sampled ibises with active infection, followed by Summer 2013, Spring 2017, and Spring 2014.
-
-```{r}
-IbisC %>%
-  count(BodyCondScore, Age, wt = HaeParasitPA)
-```
-Looking at body condition score, the majority of the birds sampled with active infection were an ideal weight or underweight.
-This lets us know that an active infection didn't affect body condition too harshly on these birds.
-
-### Figures
+#### Figures
 
@@
 ```{r}
-ExpFigure1 <-ggplot(IbisC) +
-  geom_point(aes(x = Date, y = HaeParasitLog10, color = HabType)) +
-  ggtitle("Haemoproteus parasitemia from sampled infected Ibises from 2010-2017 by habitat type") +
-  xlab("Date") +
-  ylab("Haemoproteus Parasitemia (Log base 10)") +
-  scale_color_discrete("Habitat Type", na.translate = F) +
-  theme(plot.title = element_text(hjust = 0.5, size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        legend.title = element_text(size = 10))
-
-Figure_file1 <- here("results", "Figures", "ExpFigure1.png")
-ggsave(filename = Figure_file1,plot = ExpFigure1)
-
-ExpFigure1
+knitr::include_graphics(figure_Exp_1, error = FALSE)
 ```
@@
 Overall, there is a general narrowing of the data over time with a slight negative correlation.
-There isn't a clear difference between natural and urban habitat types when it comes to amount of parasitemia.
+There isn't a clear difference between natural and urban habitat types when it comes to amount of parasitemia in the later years.
 
-Figure 2: Haemoproteus parasitemia from adult and juvenile sampled Ibises
+Figure 2: Haemoproteus parasitemia by total female and male Ibis mass
 ```{r}
-ExpFigure2 <- IbisC %>%
-  drop_na(Age) %>%
-  ggplot() +
-  geom_violin(aes(x = Age, y = HaeParasit)) +
-  ggtitle("Haemoproteus parasitemia from adult and juvenile sampled Ibises") +
-  xlab("Age") +
-  ylab("Haemoproteus Parasitemia") +
-  scale_y_continuous(limits = c(0, 0.1)) +
-  theme(plot.title = element_text(hjust = 0.5, size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        legend.title = element_text(size = 10))
-
-Figure_file2 <- here("results", "Figures", "ExpFigure2.png")
-ggsave(filename = Figure_file2, plot = ExpFigure2)
-
-ExpFigure2
-```
-
-This violin plot graphs the haemoproteus parasitemia distribution for the adult and juvenile Ibises sampled.
-The majority of the parasitemia values are on the lower end for both adults and juveniles.
-The distribution for adults reaches higher than juveniles.
-The overall structure for the adult distribution is higher and skinnier, while juveniles have more of their parasitemias lower in value.
-
-Figure 3: Haemoproteus parasitemia from sampled Ibises by sex
-```{r}
-ExpFigure3 <- IbisC %>%
-  drop_na(Sex) %>%
-  ggplot() +
-  geom_boxplot(aes(x = Sex, y = HaeParasit)) +
-  scale_y_continuous(limits = c(0, 0.5)) +
-  ggtitle("Haemoproteus parasitemia from sampled Ibises by sex") +
-  xlab("Sex") +
-  ylab("Haemoproteus Parasitemia") +
-  theme(plot.title = element_text(hjust = 0.5, size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        legend.title = element_text(size = 10))
-
-Figure_file3 <- here("results", "Figures", "ExpFigure3.png")
-ggsave(filename = Figure_file3, plot = ExpFigure3)
-
-ExpFigure3
-```
-
-The haemoproteus parasitemia distributions between sexes is extremely similar.
-The median value for males is slightly higher than females.
-The majority of the measured parasitemias are between 0.0 and 0.1.
-
-Figure 4: Haemoproteus parasitemia by total female and male Ibis mass
-```{r}
-ExpFigure4 <- IbisC %>%
-  drop_na(BirdMassG, Sex) %>%
-  ggplot() +
-  geom_point(aes(x = BirdMassG, y = HaeParasitLog10, color = Sex)) +
-  theme(axis.text.x = element_text(angle = 90)) +
-  scale_x_continuous(limits = c(130, 1240)) +
-  ggtitle("Haemoproteus parasitemia by total female and male Ibis mass") +
-  xlab("Total mass of Ibis (g)") +
-  ylab("Haemoproteus Parasitemia (Log base 10)") +
-  theme(plot.title = element_text(hjust = 0.5, size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        legend.title = element_text(size = 10))
-
-Figure_file4 <- here("results", "Figures", "ExpFigure4.png")
-ggsave(filename = Figure_file4, plot = ExpFigure4)
-
-ExpFigure4
+knitr::include_graphics(figure_Exp_4, error = FALSE)
 ```
@@
 
-Figure 5: Haemoproteus parasitemia by female and male Ibis mass and body condition score
+Figure 3: Haemoproteus parasitemia by sample site and habitat type
 ```{r}
-ExpFigure5 <- IbisC %>%
-  drop_na(BirdMassG, BodyCondScore, Sex) %>%
-  ggplot() +
-  geom_point(aes(x = BirdMassG, y = HaeParasitLog10, color = BodyCondScore)) +
-  facet_wrap(~ Sex) +
-  ggtitle("Haemoproteus parasitemia by female and male Ibis mass and body condition score") +
-  xlab("Total mass of Ibis (g)") +
-  ylab("Haemoproteus Parasitemia (Log base 10)") +
-  theme(axis.text.x = element_text(angle = 90),
-        plot.title = element_text(hjust = 0.5, size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        legend.title = element_text(size = 10)) +
-  scale_x_continuous(limits = c(130, 1240)) +
-  scale_color_discrete("Body Condition", na.translate = F)
-
-Figure_file5 <- here("results", "Figures", "ExpFigure5.png")
-ggsave(filename = Figure_file5, plot = ExpFigure5)
-
-ExpFigure5
-```
-
-Looking at this plot, there are more underweight female ibises with higher parasitemias compared to males. This could be due to the higher number of sampled females though.
-
-Figure 6: Haemoproteus parasitemia by sample site and habitat type
-```{r}
-ExpFigure6 <- IbisC %>%
-  drop_na(HabType, Site) %>%
-  ggplot() +
-  geom_point(aes(x = Site, y = log10(HaeParasit), color = HabType)) +
-  theme(axis.text.x = element_text(angle = 90),
-        plot.title = element_text(hjust = 0.5, size = 10),
-        axis.title.x = element_text(size = 10),
-        axis.title.y = element_text(size = 10),
-        legend.title = element_text(size = 10)) +
-  ggtitle("Haemoproteus parasitemia by sample site and habitat type") +
-  xlab("Sample Site") +
-  ylab("Haemoproteus Parasitemia (Log base 10)") +
-  scale_color_discrete("Habitat Type", na.translate = F)
-
-Figure_file6 <- here("results", "Figures", "ExpFigure6.png")
-ggsave(filename = Figure_file6, plot = ExpFigure6)
-
-ExpFigure6
+knitr::include_graphics(figure_Exp_6, error = FALSE)
 ```
@@
 
-Overall, there are no huge patterns or clear differences between covariates and variates so far.
-Statistical analyses may pick up on any significant differences later on, but I will return to the exploratory analysis as I continue with the project.
-I will also be meeting with those involved with the project to discuss as well.
+Overall, there are no huge patterns or clear differences between covariates and variates so far. Statistical analyses may pick up on any significant differences later on.
 
-## Statistical Analyses
+### Statistical Analyses
 
-### Presence/Absence Haemoproteus Parasitemia Data
+#### Univariate and Multivariate Logistic Models
 
@@
 ```{r}
-data_locationPA1 <- here::here("results", "Tables", "Model1MetricsPA.Rda")
-table1 <- readRDS(data_locationPA1)
-table1
+model1
 ```
-This univariate classification GLM looked at the effect of habitat type on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+This univariate classification GLM looked at the effect of age on the presence of Haemoproteus in infected birds. The results were unsatisfactory with an accuracy of 0.569 and roc_auc of 0.444.
 
 ```{r}
-data_locationPA2 <- here::here("results", "Tables", "Model2MetricsPA.Rda")
-table2 <- readRDS(data_locationPA2)
-table2
+model2
 ```
-This univariate classification GLM looked at the effect of sex of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+This univariate classification GLM looked at the effect of sex of the Ibis on the presence of Haemoproteus in infected birds.
+The results were unsatisfactory with an accuracy of 0.606 and roc_auc of 0.533, the metrics were higher than age.
 
 ```{r}
-data_locationPA3 <- here::here("results", "Tables", "Model3MetricsPA.Rda")
-table3 <- readRDS(data_locationPA3)
-table3
+model3
 ```
-This univariate classification GLM looked at the effect of age of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+This univariate classification GLM looked at the effect of sampling site of the Ibis on the presence of Haemoproteus in infected birds.
+The results were unsatisfactory with an accuracy of 0.583 and roc_auc of 0.593.
 
 ```{r}
-data_locationPA4 <- here::here("results", "Tables", "Model4MetricsPA.Rda")
-table4 <- readRDS(data_locationPA4)
-table4
+model4
 ```
-This univariate classification GLM looked at the effect of sampling site on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+This univariate classification GLM looked at the effect of sampling season on the presence of Haemoproteus in infected birds.
+The results were somewhat satisfactory with an accuracy of 0.639 and roc_auc of 0.655.
 
 ```{r}
-data_locationPA5 <- here::here("results", "Tables", "Model5MetricsPA.Rda")
-table5 <- readRDS(data_locationPA5)
-table5
+model5
 ```
-This univariate classification GLM looked at the effect of sampling date on the presence of Haemoproteus in infected birds. The results were unsatisfactory.
+This univariate classification GLM looked at the effect of sampling date on the presence of Haemoproteus in infected birds.
+The results were unsatisfactory with an accuracy of 0.539 and roc_auc of 0.548.
 
 ```{r}
-data_locationPA6 <- here::here("results", "Tables", "Model6MetricsPA.Rda")
-table6 <- readRDS(data_locationPA6)
-table6
+model6
 ```
-This univariate classification GLM looked at the effect of sampling season on the presence of Haemoproteus in infected birds.
+This univariate classification GLM looked at the effect of habitat type on the presence of Haemoproteus in infected birds.
 The results were unsatisfactory, however, these results were better than the other univariate models so far.
@@
 ```{r}
-data_locationPA7 <- here::here("results", "Tables", "Model7MetricsPA.Rda")
-table7 <- readRDS(data_locationPA7)
-table7
+model7
 ```
-This multivariate classification GLM looked at the effect of age and sex on the presence of Haemoproteus in infected birds.
-The results were unsatisfactory, this model scored about the same as the previous univariate models for age and sex.
+This univariate classification GLM looked at the effect of total bird mass in grams on the presence of Haemoproteus in infected birds.
+The results were unsatisfactory with an accuracy of 0.626 and roc_auc of 0.511.
 
 ```{r}
-data_locationPA8 <- here::here("results", "Tables", "Model8MetricsPA.Rda")
-table8 <- readRDS(data_locationPA8)
-table8
+model8
 ```
-This multivariate classification GLM looked at the effect of sampling date and sampling season on the presence of Haemoproteus in infected birds.
+This multivariate classification logistic model looked at the effect of sampling date and sampling season on the presence of Haemoproteus in infected birds.
 The results here had the highest performance of all the other models so far.
@@
 
-### Regression Models for the Amount of Haemoproteus Parasitemia
+```{r}
+model9
+```
+This multivariate classification logistic model looked at the effect of sampling date and site on the presence of *Haemoproteus* in infected birds.
+The results were unsatisfactory, just a little lower than the previous multivariate classification model.
+Sampling date and site are important variables to determine the presence of *Haemoproteus*, although season has an effect as well.
 
-This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data looking at the amount of measured parasitemia per sampled bird (on a logarithm base 10 scale), as opposed to presence/absence data.
+#### Decision Tree Model
 
-The following models were setup and tested through a 10-fold cross-validation.
-The training/testing data proportions were 80/20.
-10 folds were taken on 80% of the observations, and the results are depicted here.
-I am currently struggling with applying the models generated by CV to the testing data.
+The decision tree model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data.
+The model was run on all 10 folds of the training data.
+The roc_auc of the best tree was 0.681 and the accuracy of the model was 0.631.
+The two predictors that had the most significant effect on the accuracy of the model were the sample date and mass of the sampled birds in grams.
+The mos tree with the best parameters had a maximum depth of 8 nodes.
+
+```{r ImpPlotDT.jpeg, echo = FALSE, out.width = '100%'}
+knitr::include_graphics(figure_ML_1, error = FALSE)
+```
+Looking at the variable importance plot, we can see that sampling date and the mass of the birds in grams are the most significant predictors for the presence of *Haemoproteus* in infected birds.
+Sampling season is also important, as well as the SWA site.
+SWA had the highest number of sampled birds.
+
+```{r}
+knitr::include_graphics(figure_ML_2, error = FALSE)
+```
+This is the decision tree layout.
+Here we can see that sampling date is the primary predictor to start the tree.
+Bird mass, body condition, and SWA site are the other predictors that determine presence.
+
+#### LASSO Model
+
+I conducted a Least Absolute Shrinkage and Selection Operator (LASSO) Model with the presence or absence of *Haemoproteus* parasitemia as the outcome variable and using all the predictor variables.
+The model had an roc_auc of 0.674.
+Below is the graph of the regularization.
+
+```{r}
+knitr::include_graphics(figure_ML_3, error = FALSE)
+```
 
@@
 
-The following are the results from a random forest model run on all the predictors against the Haemoproteus parasitemia (log base 10) data.
-The missing data values were imputed by K nearest neighbors.
-The model was run on all 10 folds from the training dataset.
-```{r}
-data_locationRF <- here::here("results", "Tables", "RandomForestMetrics.Rda")
-table9 <- readRDS(data_locationRF)
-table9
+The random forest model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data. The best model resulted in an roc_auc of 0.712, which was the best so far.
+
+```{r Variable Importance Model}
+knitr::include_graphics(figure_ML_4, error = FALSE)
 ```
-This random forest model looked at the effect of all predictors on the amount of Haemoproteus in infected birds.
-The results were unsatisfactory, this model had a relatively high RMSE and low R squared value.
+Looking at the variable importance plot, it is similar to the decision tree that sample date and bird mass in grams were the top predictors.
 
-#### Linear Regression Model
+The best random forest model was run on the testing dataset, and resulted in an roc_auc of 0.725 and accuracy of 0.715. The ROC curve for that random forest model is shown below.
 
-The following are the results from a linear regression model run on habitat type against the Haemoproteus parasitemia (log base 10) data.
+```{r ROC Curve for Random Forest Model, echo = FALSE, out.width = '100%'}
+knitr::include_graphics(figure_ML_5, error = FALSE)
+```
 
-```{r}
-data_locationLM1 <- here::here("results", "Tables", "LinMetrics1.Rda")
-table10 <- readRDS(data_locationLM1)
-table10
-```
-This linear regression model looked at the effect of habitat type on the amount of Haemoproteus in infected birds.
-The results were unsatisfactory, the RMSE is high and the R squared value is very low.
+### Conclusions
 
-#### Conclusions so far
-
-This concludes the statistical analysis for part 3.
-Overall, the models did not perform well, especially those looking at the amount of Haemoproteus in infected birds.
-The presence/absence data has more potential for conclusions to be made based on generated models.
-The next steps would be to fine-tune the classification models for the presence/absence data and create a random forest model.
-
-Current problems that have arose in this process include imputing missing values to run models (as there is a large amount of missing data in this dataset), and visualizing the tidymodels results on a graph.
+This concludes the statistical analysis for part 4.
+Overall, the majority of the univariate models didn't perform well on the training dataset.
+Adding multivariate models increased the accuracy of the models but still was not accurate enough to conclude from.
+The random forest model performed the best and increased performance on the testing dataset.
+A conclusion we can make from this model is that the mass of the birds is important in determining *Haemoproteus* presence or absence, but sampling date is the most important.
+*Haemoproteus* presence in these Ibis is increasing in southern Florida regardless of sampling site.

### 2021-11-13 analysis_scripts/StatisticalAnalysisModels.R

@@ -232,7 +232,7 @@ saveRDS(Model7MetricsPA, file = "results/Tables/Model7MetricsPA.Rda")
 ##########################Multivariate GLM regression models#############################
 
 ######Model 8
-#Recipe setup for model, Parasitemia P/A vs Age and Sex on training data
+#Recipe setup for model, Parasitemia P/A vs Season and Date on training data
 PA_rec8_glm <- recipe(HaeParasitPA ~ Season + Date, data = PATrainDF)
 
 #Workflow Setup 7
@@ -418,7 +418,7 @@ saveRDS(ImpPlotRF, file = "results/Figures/ImpPlotRF.Rda")
 #Getting the predictions and residuals
 
 rf_residpredict <- final_rf_fit %>%
-  augment(new_data = PATrainDF) %>%
+  augment(PATrainDF) %>%
   select(.pred_1, HaeParasitPA)
 
 #Plot for Predicted vs Observed Values
@@ -451,8 +451,11 @@ ROC_Test_RF <- Test_rf_fit %>%
   collect_predictions() %>% 
   roc_curve(HaeParasitPA, .pred_0) %>% 
   autoplot()
+
 saveRDS(ROC_Test_RF, file = "results/Figures/RF_Test_ROCPlot.Rda")
 
+file_location <- here("results", "Figures", "ROC_TestPlot")
+ggsave(filename = file_location, plot = ROC_Test_RF)
 
 
 

### 2021-11-28 products/manuscript/Manuscript.Rmd
Finished Part 5 11/28/21
Part 5 of the project is completed and is ready for peer-review.

@@
 ---
-title: "Prevalence of *Haemoproteus plataleae* in American White Ibis from Southern Florida"
+title: "Prevalence of *Haemoproteus* parasites in American White Ibis from Southern Florida"
 subtitle: ""
@@
 ---
+
+## Abstract
+
+Blood parasites of the *Haemoproteus* genus can infect various bird species and cause chronic infections that last an entire lifetime.
+A general increase in the presence of parasitic pathogens of birds led to the collection of blood samples from American White Ibis in Southern Florida from 2010-2017.
+A data analysis on the demographic, temporal, spatial, and parasitemia data collected from over 300 sampled Ibis was conducted.
+Various univariate and multivariate logistic regression models, along with a random forest model, concluded that the sampling date was the most important predictor in determining whether a sampled Ibis had a *Haemoproteus* infection (ROC-AUC 0.725, Accuracy 0.715).
+This led to the observation that more Ibis have *Haemoproteus* infections in Southern Florida in 2017 compared to 2010, however, there are numerous other epidemiologic and pathogenic factors about the parasite that makes any further conclusions difficult.
+An increased sampling effort and further studies are needed.
+
+## Loading Required Data and Packages
 ```{r}
@@
 library(tidyverse) #all required data manipulation packages
-library(xlsx) #to read in excel spreadsheets
-library(janitor) #To assist in cleaning up dates and tidy data
-library(lubridate) #To extract times from raw data
 library(tidymodels) #to run the statistical analyses
@@
 #Machine Learning Model Diagnostic Figures
-figure_ML_1 <- here::here("results", "Figures", "ImpPlotDT.jpeg")
-figure_ML_2 <- here::here("results", "Figures", "DTTree.jpeg")
-figure_ML_3 <- here::here("results", "Figures", "LassoFig.jpeg")
 figure_ML_4 <- here::here("results", "Figures", "ImpPlotRF.jpeg")
@@
 #Loading Model Results
-data_locationPA1 <- here::here("results", "Tables", "Model1MetricsPA.Rda")
-model1 <- readRDS(data_locationPA1)
-data_locationPA2 <- here::here("results", "Tables", "Model2MetricsPA.Rda")
-model2 <- readRDS(data_locationPA2)
-data_locationPA3 <- here::here("results", "Tables", "Model3MetricsPA.Rda")
-model3 <- readRDS(data_locationPA3)
 data_locationPA4 <- here::here("results", "Tables", "Model4MetricsPA.Rda")
@@
 model5 <- readRDS(data_locationPA5)
-data_locationPA6 <- here::here("results", "Tables", "Model6MetricsPA.Rda")
-model6 <- readRDS(data_locationPA6)
-data_locationPA7 <- here::here("results", "Tables", "Model7MetricsPA.Rda")
-model7 <- readRDS(data_locationPA7)
 data_locationPA8 <- here::here("results", "Tables", "Model8MetricsPA.Rda")
 model8 <- readRDS(data_locationPA8)
-data_locationPA9 <- here::here("results", "Tables", "Model9MetricsPA.Rda")
-model9 <- readRDS(data_locationPA9)
-#Loading required materials - Tables
-mod_loc_1 <- here::here("results", "Models", "Model1PA.Rda")
-mod_loc_2 <- here::here("results", "Models", "Model2PA.Rda")
-mod_loc_3 <- here::here("results", "Models", "Model3PA.Rda")
-mod_loc_4 <- here::here("results", "Models", "Model4PA.Rda")
-mod_loc_5 <- here::here("results", "Models", "Model5PA.Rda")
-mod_loc_6 <- here::here("results", "Models", "Model6PA.Rda")
-mod_loc_7 <- here::here("results", "Models", "Model7PA.Rda")
-mod_loc_8 <- here::here("results", "Models", "Model8PA.Rda")
-mod_loc_9 <- here::here("results", "Models", "Model9PA.Rda")
-
 ```
@@
 
-## General Background Information
+Avian haemosporidians are vector-borne parasites of the blood that have a global distribution [@Clark2014].
+*Plasmodium* and *Haemoproteus* are two of the most common genera, however, *Haemoproteus* is much more common among avian hosts [@Clark2014].
+They are believed to be the least pathogenic and have a narrow host range [@Atkinson1991].
+Previous studies have shown that *Haemoproteus* parasitemia has been detected in Green Herons (*Butorides virescens*), Roseate Spoonbills (*Platalea ajaja*), Glossy Ibis (*Plegadis falcinellus*), and American White Ibis (*Eudocimus albus*) [@Coker2017].
+
+### General Background Information
 The data this analysis concerns is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017.
@@
 Haemosporidian infections usually cause long-lasting chronic infections in natural hosts, and can cause reduced reproductive success, host fitness, and increased stress and disease susceptibility [@Coker2017].
+Lower parasitemia values suggest that the infections in the Ibis are chronic [@Coker2017].
 
-The purpose of this analysis is to determine whether the presence of *Haemoproteus* was increasing over time in Ibis in southern Florida, and if so, determine any demographic trends among the Ibis or temporal or spatial trends among the system at large.
-Here we investigate whether age, sex, sampling site, capture season, date of capture, body condition, and/or mass of captured Ibis make significant differences on the overall presence of *Haemoproteus* in the birds.
+The purpose of this analysis is to determine whether the presence of *Haemoproteus* was increasing over time in Ibis in southern Florida, and if so, determine any demographic trends among the Ibis, or temporal or spatial trends among *Haemoproteus* ecology.
+Here we investigate whether age, sex, sampling site, capture season, date of capture, body condition, and mass of captured Ibis make significant differences on the overall presence of *Haemoproteus* in the birds.
 
@@
 20 Ibis were removed from the analysis because they did not have a positive *Haemoproteus* parasitemia value and the site from which they were captured had less than or equal to 5 sampled Ibis and had no presence of infected Ibis.
-Ibis were either classified as having parasitemia (1 - presence) or not having parasitemia (0 - absence).
+Ibis were either classified as having parasitemia (1 = presence) or not having parasitemia (0 = absence).
 
@@
 The following tables and figures show those relationships prior to the statistical analysis.
+Tables and figures not included here are located in the supplemental file.
 
@@
 ```
-Overall, looking at the distribution of Ibis Sex and Age, the are more females than males and adults than juveniles.
+Overall, looking at the distribution of Ibis sex and age, the are more females than males and adults than juveniles.
 Female adults have the highest number with 109.
@@
 This will most likely have to be completed with and without imputation depending on the model used.
+The results could be somewhat biased due to this distribution.
 
@@
 ```
-This table shows the number of Ibis sampled during each season. The majority of the Ibis were sampled in the later years (2015-2017) than in the earlier years (2010-2014).
+This table shows the number of Ibis sampled during each season.
+The majority of the Ibis were sampled in the later years (2015-2017) than in the earlier years (2010-2014).
+13 samples did not have a recorded season.
+The highest number of Ibis sampled were in Spring, followed by Summer in the earlier years and Fall in the later years.
 
@@
 ```
-The majority of the Ibis were sampled in SWA, DRP, and ICP sites. Most of the Ibis were sampled in urban sites compared to natural sites.
+The majority of the Ibis were sampled in SWA, DRP, and ICP. Some sites had very low sampling numbers. Most of the Ibis were sampled in urban sites compared to natural sites.
 
@@
 Now we can move onto creating some figures for the data to see any patterns visually.
+Haemoproteus parasitemia was transformed on a log base-10 scale for the exploratory figures here.
+This was conducted in order to see if there was any clear patterns in the amount of parasitemia among Ibis with *Haemoproteus* infection.
 
@@
 
-Looking at the plot, male birds seem to weigh more than females. However, when comparing the parasitemia between females and males, there isn't a clear difference between the two.
+Looking at the plot, male birds seem to weigh more than females.
+However, when comparing the parasitemia between females and males, there isn't a clear difference between the two.
+There seems to be an hourglass shape in the graph however, with majority of the parasitemia values either being a higher or lower value.
 
@@
 
-Overall, there are no huge patterns or clear differences between covariates and variates so far. Statistical analyses may pick up on any significant differences later on.
+Overall, there are no huge patterns or clear differences between covariates and variates so far.
 
@@
 
-This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data just looking at the presence and absence data, as opposed to the amount of parasitemia per sampled bird.
+This section details the results of the statistical analysis on the Haemoproteus parasitemia data just looking at the presence and absence data, as opposed to the amount of parasitemia per sampled bird.
 
-The following models were setup and tested through a 10-fold cross-validation.
+The following models were setup and tested through 10-fold cross-validation.
 The training/testing data proportions were 80/20.
 10 folds were taken on 80% of the observations, and the results are depicted here.
-I am currently struggling with applying the models generated by CV to the testing data.
-
-```{r}
-model1
-```
-This univariate classification GLM looked at the effect of age on the presence of Haemoproteus in infected birds. The results were unsatisfactory with an accuracy of 0.569 and roc_auc of 0.444.
-
-```{r}
-model2
-```
-This univariate classification GLM looked at the effect of sex of the Ibis on the presence of Haemoproteus in infected birds.
-The results were unsatisfactory with an accuracy of 0.606 and roc_auc of 0.533, the metrics were higher than age.
-
-```{r}
-model3
-```
-This univariate classification GLM looked at the effect of sampling site of the Ibis on the presence of Haemoproteus in infected birds.
-The results were unsatisfactory with an accuracy of 0.583 and roc_auc of 0.593.
+Models with variables not included here are included in the supplemental information.
 
@@
 ```
-This univariate classification GLM looked at the effect of sampling season on the presence of Haemoproteus in infected birds.
+This univariate logistic regression looked at the effect of sampling season on the presence of *Haemoproteus* in infected birds.
 The results were somewhat satisfactory with an accuracy of 0.639 and roc_auc of 0.655.
@@
 ```
-This univariate classification GLM looked at the effect of sampling date on the presence of Haemoproteus in infected birds.
-The results were unsatisfactory with an accuracy of 0.539 and roc_auc of 0.548.
-
-```{r}
-model6
-```
-This univariate classification GLM looked at the effect of habitat type on the presence of Haemoproteus in infected birds.
-The results were unsatisfactory, however, these results were better than the other univariate models so far.
-
-```{r}
-model7
-```
-This univariate classification GLM looked at the effect of total bird mass in grams on the presence of Haemoproteus in infected birds.
-The results were unsatisfactory with an accuracy of 0.626 and roc_auc of 0.511.
+This univariate logistic regression looked at the effect of sampling date on the presence of *Haemoproteus* in infected birds.
+The results were somewhat unsatisfactory with an accuracy of 0.539 and roc_auc of 0.548.
 
@@
 ```
-This multivariate classification logistic model looked at the effect of sampling date and sampling season on the presence of Haemoproteus in infected birds.
+This multivariate logistic regression model looked at the effect of sampling date and sampling season on the presence of Haemoproteus in infected birds.
 The results here had the highest performance of all the other models so far.
-One way to interpret this model would be that as the sampling effort continued from 2010-2017, the passage of time and season had a significant effect on the presence of Haemoproteus in Ibis.
-However, the performance of this model isn't high enough to successfully conclude with that explanation.
+As the sampling effort continued from 2010-2017, the passage of time and season had a significant effect on the presence of *Haemoproteus* in Ibis.
+However, the performance of this model isn't high enough to justify successfully concluding with that explanation.
 
-```{r}
-model9
-```
-This multivariate classification logistic model looked at the effect of sampling date and site on the presence of *Haemoproteus* in infected birds.
-The results were unsatisfactory, just a little lower than the previous multivariate classification model.
-Sampling date and site are important variables to determine the presence of *Haemoproteus*, although season has an effect as well.
+### Random Forest Model
 
-#### Decision Tree Model
-
-The decision tree model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data.
-The model was run on all 10 folds of the training data.
-The roc_auc of the best tree was 0.681 and the accuracy of the model was 0.631.
-The two predictors that had the most significant effect on the accuracy of the model were the sample date and mass of the sampled birds in grams.
-The mos tree with the best parameters had a maximum depth of 8 nodes.
-
-```{r ImpPlotDT.jpeg, echo = FALSE, out.width = '100%'}
-knitr::include_graphics(figure_ML_1, error = FALSE)
-```
-Looking at the variable importance plot, we can see that sampling date and the mass of the birds in grams are the most significant predictors for the presence of *Haemoproteus* in infected birds.
-Sampling season is also important, as well as the SWA site.
-SWA had the highest number of sampled birds.
-
-```{r}
-knitr::include_graphics(figure_ML_2, error = FALSE)
-```
-This is the decision tree layout.
-Here we can see that sampling date is the primary predictor to start the tree.
-Bird mass, body condition, and SWA site are the other predictors that determine presence.
-
-#### LASSO Model
-
-I conducted a Least Absolute Shrinkage and Selection Operator (LASSO) Model with the presence or absence of *Haemoproteus* parasitemia as the outcome variable and using all the predictor variables.
-The model had an roc_auc of 0.674.
-Below is the graph of the regularization.
-
-```{r}
-knitr::include_graphics(figure_ML_3, error = FALSE)
-```
-
-#### Random Forest Model
-
-The random forest model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data. The best model resulted in an roc_auc of 0.712, which was the best so far.
+A random forest model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data.
+The number of features to consider at each split and the minimum number of data points required for the node to be split were both tuned with a tuning grid.
+The random forest model consisted of 500 different decision trees.
+The best model resulted in an roc_auc of 0.712 on the training data.
 
@@
 ```
-Looking at the variable importance plot, it is similar to the decision tree that sample date and bird mass in grams were the top predictors.
+The variable importance plot suggests that the sampling date and mass of the sampled bird in grams are the two most important predictors of *Haemoproteus* presence in the sampled Ibis.
 
-The best random forest model was run on the testing dataset, and resulted in an roc_auc of 0.725 and accuracy of 0.715. The ROC curve for that random forest model is shown below.
+The most accurate random forest model was run on the testing dataset, which resulted in an roc_auc of 0.725 and accuracy of 0.715. The ROC curve for that random forest model is shown below.
 
@@
 
+### Discussion
+
+Based on the analysis conducted on the *Haemoproteus* dataset, sampling date and mass of the sampled Ibis are the most important predictors in determining whether the sampled Ibis will develop parasitemia.
+This study has shown that the number of *Haemoproteus* infections in Ibis are increasing in southern Florida from 2010 to 2017.
+The mass of the sampled Ibis is also an important predictor in determining whether or not the Ibis will develop an infection.
+
+However, there are several other determinants in the *Haemoproteus* epidemiology and pathogenesis that can influence the development of infection.
+The results from this analysis aren't powerful enough to successfully conclude that the mass of the birds is a good predictor.
+The life cycle of the *Culicoides* midges, pathogenesis of the *Haemoproteus* parasite itself in both the vector and host organism, and *Culicoides* biting rate could all have a more significant effect but can't be directly applied to this dataset.
+
+It may also be the case that the results of the analysis suggesting sampling date as the most important predictor is simply due to the higher number of Ibis being sampled in the later years of the study.
+The other individual predictors were all used in univariate logistic regression models, and none of them had highly significant results.
+
 ### Conclusions
 
-This concludes the statistical analysis for part 4.
-Overall, the majority of the univariate models didn't perform well on the training dataset.
-Adding multivariate models increased the accuracy of the models but still was not accurate enough to conclude from.
-The random forest model performed the best and increased performance on the testing dataset.
-A conclusion we can make from this model is that the mass of the birds is important in determining *Haemoproteus* presence or absence, but sampling date is the most important.
-*Haemoproteus* presence in these Ibis is increasing in southern Florida regardless of sampling site.
+This concludes the statistical analysis for the *Haemoproteus* presence and absence data.
+Overall, there isn't a clear predictor variable that determines whether or not sampled Ibis will have *Haemoproteus* parasitemia.
+The random forest model performed the best and had increased performance on the testing dataset.
+It suggested that sampling date was the most important predictor in determining whether the sampled Ibis would develop *Haemoproteus* parasitemia or not, but this may just be the case of their being a higher number of sampled Ibis in the later years.
+However, there are numerous other factors regarding the *Haemoproteus* epidemiology and pathogenesis that make any further conclusions difficult.

### 2021-11-28 supplemental_materials/SupplementalFile.Rmd
Finished Part 5 11/28/21
Part 5 of the project is completed and is ready for peer-review.

---
title: "Supplemental File"
author: "Ryan Grunert"
date: "11/26/2021"
output: word_document
---

```{r}
#Loading required packages
library(here)
library(tidyverse)
library(knitr)
#Exploratory Figures
figure_Exp_2 <- here::here("supplemental_materials", "Figures", "ExpFigure2.png")
figure_Exp_3 <- here::here("supplemental_materials", "Figures", "ExpFigure3.png")
figure_Exp_5 <- here::here("supplemental_materials", "Figures", "ExpFigure5.png")
#Machine Learning Model Diagnostic Figures
figure_ML_1 <- here::here("supplemental_materials", "Figures", "ImpPlotDT.jpeg")
figure_ML_2 <- here::here("supplemental_materials", "Figures", "DTTree.jpeg")
figure_ML_3 <- here::here("supplemental_materials", "Figures", "LassoFig.jpeg")
#Loading Model Results
data_locationPA1 <- here::here("supplemental_materials", "Tables", "Model1MetricsPA.Rda")
model1 <- readRDS(data_locationPA1)
data_locationPA2 <- here::here("supplemental_materials", "Tables", "Model2MetricsPA.Rda")
model2 <- readRDS(data_locationPA2)
data_locationPA3 <- here::here("supplemental_materials", "Tables", "Model3MetricsPA.Rda")
model3 <- readRDS(data_locationPA3)
data_locationPA6 <- here::here("supplemental_materials", "Tables", "Model6MetricsPA.Rda")
model6 <- readRDS(data_locationPA6)
data_locationPA7 <- here::here("supplemental_materials", "Tables", "Model7MetricsPA.Rda")
model7 <- readRDS(data_locationPA7)
data_locationPA9 <- here::here("supplemental_materials", "Tables", "Model9MetricsPA.Rda")
model9 <- readRDS(data_locationPA9)
#Loading required materials - Tables
mod_loc_1 <- here::here("results", "Models", "Model1PA.Rda")
mod_loc_2 <- here::here("results", "Models", "Model2PA.Rda")
mod_loc_3 <- here::here("results", "Models", "Model3PA.Rda")
mod_loc_4 <- here::here("results", "Models", "Model4PA.Rda")
mod_loc_5 <- here::here("results", "Models", "Model5PA.Rda")
mod_loc_6 <- here::here("results", "Models", "Model6PA.Rda")
mod_loc_7 <- here::here("results", "Models", "Model7PA.Rda")
mod_loc_8 <- here::here("results", "Models", "Model8PA.Rda")
mod_loc_9 <- here::here("results", "Models", "Model9PA.Rda")
```

This document details all of the supplemental information and models generated by the project.

### Exploratory Data Analysis

The following are exploratory data analysis graphs that were considered vital enough to include in the main manuscript.

```{r}
knitr::include_graphics(figure_Exp_2, error = FALSE)
```
This is a violin plot of the *Haemoproteus* parasitemia values among infected adult and juvenile white Ibis. As we can see, the majority of the birds have lower parasitemias below 0.025, and a couple outliers that have higher values (most likely due to recent infection, not chronic). The adult birds have a slightly larger spread than the juveniles.

```{r}
knitr::include_graphics(figure_Exp_3, error = FALSE)
```
This figure is a boxplot of the *Haemoproteus* parasitemia distributions of female and male infected Ibis. The females have a slightly larger distribution, but no other clear pattern exists between the sexes.

```{r}
knitr::include_graphics(figure_Exp_5, error = FALSE)
```
This figure is a scatter plot of *Haemoproteus* parasitemia values and individual Ibis body condition stratified by sex. While the male birds do have larger masses, the overall distribution of body conditions is similar between males and females regardless of parasitemia value. This supports the theory that chronic infections with the parasite don't have a measurable detrimental health effect on the Ibis themselves.

### Univariate Logistic Models

This section details the results of a preliminary statistical analysis on the Haemoproteus parasitemia data just looking at the presence and absence data, as opposed to the amount of parasitemia per sampled bird.

The following models were setup and tested through a 10-fold cross-validation. The training/testing data proportions were 80/20. 10 folds were taken on 80% of the observations, and the results are depicted here.

```{r}
model1
```
This univariate classification GLM looked at the effect of age on the presence of Haemoproteus in infected birds. The results were unsatisfactory with an accuracy of 0.569 and roc_auc of 0.444.

```{r}
model2
```
This univariate classification GLM looked at the effect of sex of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory with an accuracy of 0.606 and roc_auc of 0.533, the metrics were higher than age.

```{r}
model3
```
This univariate classification GLM looked at the effect of sampling site of the Ibis on the presence of Haemoproteus in infected birds. The results were unsatisfactory with an accuracy of 0.583 and roc_auc of 0.593.

```{r}
model6
```
This univariate classification GLM looked at the effect of habitat type on the presence of Haemoproteus in infected birds. The results were unsatisfactory, however, these results were better than the other univariate models so far.

```{r}
model7
```
This univariate classification GLM looked at the effect of total bird mass in grams on the presence of Haemoproteus in infected birds. The results were unsatisfactory with an accuracy of 0.626 and roc_auc of 0.511.

### Multivariate Models

```{r}
model9
```
This multivariate classification logistic model looked at the effect of sampling date and site on the presence of *Haemoproteus* in infected birds. The results were unsatisfactory, just a little lower than the previous multivariate classification model. Sampling date and site are important variables to determine the presence of *Haemoproteus*, although season has an effect as well.

### LASSO Model

I conducted a Least Absolute Shrinkage and Selection Operator (LASSO) Model with the presence or absence of *Haemoproteus* parasitemia as the outcome variable and using all the predictor variables. The model had an roc_auc of 0.674. Below is the graph of the regularization.

```{r}
knitr::include_graphics(figure_ML_3, error = FALSE)
```

#### Decision Tree Model

The decision tree model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data. The model was run on all 10 folds of the training data. The roc_auc of the best tree was 0.681 and the accuracy of the model was 0.631. The two predictors that had the most significant effect on the accuracy of the model were the sample date and mass of the sampled birds in grams. The mos tree with the best parameters had a maximum depth of 8 nodes.

```{r ImpPlotDT.jpeg, echo = FALSE, out.width = '100%'}
knitr::include_graphics(figure_ML_1, error = FALSE)
```
Looking at the variable importance plot, we can see that sampling date and the mass of the birds in grams are the most significant predictors for the presence of *Haemoproteus* in infected birds. Sampling season is also important, as well as the SWA site. SWA had the highest number of sampled birds.

This concludes the supplemental information for the project.

### 2021-11-28 analysis_scripts/ExploratoryAnalysisScript.R

@@ -166,7 +166,7 @@ ExpFigure2 <- IbisC %>%
         axis.title.y = element_text(size = 10),
         legend.title = element_text(size = 10))
 
-Figure_file2 <- here("results", "Figures", "ExpFigure2.png")
+Figure_file2 <- here("supplemental_materials", "Figures", "ExpFigure2.png")
 ggsave(filename = Figure_file2, plot = ExpFigure2)
 
 #########Plot 3: Boxplot of haemoproteus parasitemia vs sex
@@ -184,7 +184,7 @@ ExpFigure3 <- IbisC %>%
         axis.title.y = element_text(size = 10),
         legend.title = element_text(size = 10))
 
-Figure_file3 <- here("results", "Figures", "ExpFigure3.png")
+Figure_file3 <- here("supplemental_materials", "Figures", "ExpFigure3.png")
 ggsave(filename = Figure_file3, plot = ExpFigure3)
 
 ##########Plot 4: Scatterplot of haemoproteus parasitemia vs bird weight with sex
@@ -223,7 +223,7 @@ ExpFigure5 <- IbisC %>%
   scale_x_continuous(limits = c(130, 1240)) +
   scale_color_discrete("Body Condition", na.translate = F)
 
-Figure_file5 <- here("results", "Figures", "ExpFigure5.png")
+Figure_file5 <- here("supplemental_materials", "Figures", "ExpFigure5.png")
 ggsave(filename = Figure_file5, plot = ExpFigure5)
 
 ###########Plot 6: Scatterplot of haemoptroteus parasitemia vs sample site and habitat type

### 2021-11-28 analysis_scripts/StatisticalAnalysisModels.R

@@ -109,8 +109,8 @@ Model1MetricsPA <- collect_metrics(PA_fit1)
 #Accuracy = 0.545, roc_auc = 0.451
 
 #Saving the results and table
-saveRDS(PA_fit1, file = "results/Models/Model1PA.Rda")
-saveRDS(Model1MetricsPA, file = "results/Tables/Model1MetricsPA.Rda")
+saveRDS(PA_fit1, file = "supplemental_materials/Models/Model1PA.Rda")
+saveRDS(Model1MetricsPA, file = "supplemental_materials/Tables/Model1MetricsPA.Rda")
 
 
 
@@ -128,8 +128,8 @@ Model2MetricsPA <- collect_metrics(PA_fit2)
 #Accuracy = 0.606, roc_auc = 0.534
 
 #Saving the results and table
-saveRDS(PA_fit2, file = "results/Models/Model2PA.Rda")
-saveRDS(Model2MetricsPA, file = "results/Tables/Model2MetricsPA.Rda")
+saveRDS(PA_fit2, file = "supplemental_materials/Models/Model2PA.Rda")
+saveRDS(Model2MetricsPA, file = "supplemental_materials/Tables/Model2MetricsPA.Rda")
 
 
 
@@ -147,8 +147,8 @@ Model3MetricsPA <- collect_metrics(PA_fit3)
 #Accuracy = 0.584, roc_auc = 0.594
 
 #Saving the results and table
-saveRDS(PA_fit3, file = "results/Models/Model3PA.Rda")
-saveRDS(Model3MetricsPA, file = "results/Tables/Model3MetricsPA.Rda")
+saveRDS(PA_fit3, file = "supplemental_materials/Models/Model3PA.Rda")
+saveRDS(Model3MetricsPA, file = "supplemental_materials/Tables/Model3MetricsPA.Rda")
 
 
 
@@ -205,8 +205,8 @@ Model6MetricsPA <- collect_metrics(PA_fit6)
 #Accuracy = 0.535, roc_auc = 0.476
 
 #Saving the results and table
-saveRDS(PA_fit6, file = "results/Models/Model6PA.Rda")
-saveRDS(Model6MetricsPA, file = "results/Tables/Model6MetricsPA.Rda")
+saveRDS(PA_fit6, file = "supplemental_materials/Models/Model6PA.Rda")
+saveRDS(Model6MetricsPA, file = "supplemental_materials/Tables/Model6MetricsPA.Rda")
 
 
 ######Model 7
[…]

### 2021-12-03 Project Review A. Foley.Rmd
A. Foley Project Review

---
title: Project Review A. Foley 
date: "`r file.mtime(knitr::current_input())`"
#bibliography: ../media/references.bib
output: 
  html_document:
    toc_depth: 3
    number_sections: true
---

# Overview

Title of project: Prevalence of Haemoproteus parasites in American White Ibis from Southern Florida

Name of project author(s): Ryan Grunert

Name of project reviewer: Amelia Foley

# Specific project content evaluation

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

The abstract provides a succinct summary of the project and the introduction and background information elaborate an appropriate amount on the context of the study. I can understand where this research fits into the current body of work and see that this is supported by a couple of references. To add to the motivation aspect, perhaps you could include a sentence or two elaborating on why studying this is important (ecological implications, etc).

Unsure if "loading required data and packages" section is purposefully included in the manuscript - should this be set to echo=FALSE?

### Summary assessment
* strong contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

The project aims are easily found in the general background information. It's clear that the main question regard whether Haemoproteus increases over tiem in Ibis, with secondary question about demographic trends. The variables of interest are clearly listed.

### Summary assessment
* question/hypotheses fully clear

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

The source of the data is well-described. I was able to locate the meta data word doc in the products folder, but initially was looking for it in the raw data folder (I think it might make more sense to have it there). The meta data provides clear definitions for the raw data. I see that you note the number of Ibis sampled in the study population section, but I might also include the number of observations and variables earlier in the data collection section, to provide the reader a general scope of the study size.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

The cleaning process is summarized in the manuscript and also described in notes in the processing script, making it easy to follow along. I would remove the first person "I" statement from the data cleaning section and switch to third-person.

This line appears a few times in the manuscript - knitr::include_graphics(figure_Exp_1, error = FALSE)
This comes up between your figure caption and the figure, making it seem a little disconnected.

In your supplemental file, I think you might need an extra line between your figures and the text following them - currently the paragraph starts to the right of the image rather than on a new line.

The exploratory analysis process was easy to follow along and I understood your thought process through the figures and your explanations following them.

Some of your figures lack captions or figure numbers - adding these might be helpful for the organization of the supplement and manuscript

### Summary assessment
* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

Analysis is well structured and explained, both in the manuscript and the code. Multiple univariate models were performed, as well as a multivariate model and three machine learning models, which seems to really strengthen the analysis. You may want to explicitly state that you only report the random forest model in the manuscript because it was the best performing model (possibly citing the comparison you provide in the supplementary material).

Model performance metrics are clearly stated in the methods section, and continually referred to throughout the manuscript to add context to the results. Your interpretation of the analysis seems justified and well-supported by the metrics.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

Some figures are in need of captions/figure numbers so that readers can easily ID them

Figures are tables are easy to understand. The figures appear publication ready though there is not much formatting for the tables - they are readable but might require a different format to be considered "publication level quality"

### Summary assessment
* results are very well presented

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

Findings are well summarized in the first paragraph of the discussion, and limitations are addressed thoroughly. If there are particular strengths to this study (compared to previous ones, such as longer study period, greater number of observations, level of analysis), maybe you could state some strengths in this section.

Your final statement about numerous other factors regarding infection making further conclusions difficult seems a bit vague. What other factors? Maybe you could reframe this by defining specific needs for further studies. Between that sentence and the sentence prior, it seems to take value away from your conclusion. Definitely important to acknowledge limitations but maybe you can end the paragraph on a stronger note?

In discussion section, their -> there

### Summary assessment
* strong, complete and clear discussion

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

The whole repo is well organized. The file names make sense and there don't appear to be an extra or unneccesary files that don't serve a particular purpose. The main readme files had all of the info I needed to reproduce the analysis, but even without the readme the files are labelled well enough that I could probably infer the correct order/structure of things. I like that you have your results folder further organized into the different type of result (figure, model, table)

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Processing script very well organized/documented. All scripts are neat and contain a note explaining what each line of code does, very easy to follow along.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Received this error running the processing script even after installing rJava (or attempting to) - could definitely be more of an issue with my computer than with your code. Maybe the function `read_excel()` would be more accessible in terms of reproducibility, compared to `read.xlsx` from the `xlsx` packages?

library(xlsx) #to read in excel spreadsheets
The operation couldn’t be completed. Unable to locate a Java Runtime.
Please visit http://www.java.com for information on installing Java.

Error: package or namespace load failed for ‘xlsx’:
 .onLoad failed in loadNamespace() for 'rJava', details:
  call: fun(libname, pkgname)
  error: JVM could not be found
In addition: Warning messages:
1: In system("/usr/libexec/java_home", intern = TRUE) :
  running command '/usr/libexec/java_home' had status 1
2: In fun(libname, pkgname) :
  Cannot find JVM library 'NA/lib/server/libjvm.dylib'
Install Java and/or check JAVA_HOME (if in doubt, do NOT set it, it will be detected)

Exploratory analysis ran easily. Your warning/note about the run time for the statistical analysis script is appreciated! The statistical analysis script ran successfully but did take a long time, as you noted.

### Summary assessment
* small parts not reproducible or required manual intervention

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Multiple models were considered. There were minimal data processing decisions required beyond cleaning, so alternatives don't seem neccessary here. The main question about infections over time is clearly addressed in the discussion/conclusion. There isn't much mention of additional demographic variables in the discussion/conclusion, though this may be due to lack of relationship shown (represented in variable importance figure)

### Summary assessment
* strong level of thoroughness

## Further comments

The section numbers in your manuscript all begin with 0, as in 0.4.2 instead of 4.2 - it may look neater to remove the 0 if you can figure out what causes that

I think you may be missing a title for your references section (?)

Your code for this project look great! All of the scripts are super well organized and easy to follow, and it is obvious that you have a good understanding of the statistical analysis/modeling portion, it seems very well done. I think I would just consider a few small changes with those formatting things in the manuscript and a little more elaboration in the discussion and conclusion, overall this seems like a very good quality analysis

### 2021-12-03 Project_Review_Grunert.Rmd
Project Review - CC

# Overview

Title of project: "Prevalence of *Haemoproteus* parasites in American White Ibis from Southern Florida"

Name of project author(s): Ryan Grunert

Name of project reviewer: Carter Coleman

# Specific project content evaluation
Evaluate the different parts of the project by filling in the sections below.

## Background, Context and Motivation
How well is the context of the project described? Is a comprehensive background, including summary of previous/related work given? Is the project well placed into the context of existing work (including proper referencing of existing work). Is it clear why the project was undertaken and what new information it hopes to provide?

### Feedback and Comments

You did a good job describing the mechanisms of transmission of Haemoproteus in American White Ibis populations. However, you are missing the motivation for why this is important. Are the chronic issues in Ibis causing population decline in other areas? Why in South Florida? I just think a little more info and literature on why you are usung the data you are using is needed.

### Summary assessment

* some contextualization and motivation

## Question description
How well and clear are the question(s)/hypotheses the project aims to address described? Is it clear how the questions relate to the data?

### Feedback and Comments

Your question is clear. Looks like you are interested in looking at population dynamics of Haemoproteus in Southern Floridian White Ibis populations by assessing correlation of different population chamging parameters with Haemoproteus presence. However, I didn't see a hypothesis clearly stated. I would make sure you include how you expect the Haemoproteus population to change with certain important variables used.

### Summary assessment
* question/hypotheses somewhat explained

## Data description
How well is the data overall described? Is the source provided? Is a codebook or other meta-information available that makes it clear what the data is?

### Feedback and Comments

Nice and concise explanation of what is in the data set and when/where/who collected the data. No recommendations.

### Summary assessment
* source and overall structure of data well explained

## Data wrangling and exploratory analysis
How well is the data cleaned/processed and explored? Are all steps reasonable and well explained? Are alternatives discussed and considered? Are meaningful exploratory results shown (e.g. in the supplementary materials)?

### Feedback and Comments

From what I could tell, your data cleaning and exploration was very thorough. I was able to find a biological reason for every processing decision made. Additionally, I like how you incorporated the NA values into the exploration, rather than finding a reason to remove them. I liked your use of the exploration as a way to show distribution of your outcome variable across different potential predictor variables. It was a great use of box, violin, and scatter plots to see if their were any intereting distributions of Haemoproteus.

However, I was not able to run your processinf script because of issues with the "xlsx" package. I was not even able to install the package. I would check into that issue.

### Summary assessment

* essentially no weaknesses in wrangling and exploratory component

## Appropriateness of Analysis
Were the analysis methods appropriate for the data? Was the analysis done properly? Were different components of the analysis (e.g. performance measure, variable selection, data pre-processing, model evaluation) done in the best way possible and explained well?

### Feedback and Comments

The types of analyses were appropriate and thorough; Staring with univariate analysis of association and moving to three different machine learning models to assess best combination of predictor variables. However, I would include some text about which machine learning model fit the data best, based on results and model limitations.

### Summary assessment
* strong and reasonable analysis

## Presentation
How well are results presented? Are tables and figures easy to read and understand? Are the main figures/tables publication level quality?

### Feedback and Comments

I think all your figures look great! However, go back through and make sure that the code chunks are not present in you manuscript. You can do this by including the following at the start of the code chunk: {r, echo=FALSE, message=FALSE}.

### Summary assessment
* results are presented ok, with room for improvement

## Discussion/Conclusions
Are the study findings properly discussed? Are strengths and limitations acknowledged? Are findings interpreted properly?

### Feedback and Comments

I liked how you made some softer inferences on what variables may be important in the spread of Haemoproteus in the Whit Ibis populations. By indicating sampling date and mass of bird as potentially strong predictors, it points future research towards understanding seasonality of Haemoproteus spread and support of the traditional parasite hypothesis that underdevelopment and lack of proper growth (indicated by low mass for age) makes one moresuceptible to infection. I would find some literature to back those findings up in the conclusion if you are going to state them specifically.

### Summary assessment
* minor parts wrong, missing or unclear

## Further comments

_Great work! I think most of my comments were about elaborating on what you have further. Nothing is really missing and your analysis was very thorough. I just think statements could be elaborated on and backed up more throughout._

# Overall project content evaluation
Evaluate overall features of the project  by filling in the sections below.

## Structure
Is the project well structured? Are files in well labeled folders? Do files have reasonable names? Are all "junk" files not needed for analysis/reproduction removed? By just looking at files and folders, can you get an idea of how things fit together?

### Feedback and Comments

Your structure was great! very well organized and extremely easy to navigate. I barely had to use your readmes to find and run things. I would suggest breaking up parts of your statistical analyses. When you have all machine learning models in one folder, it gets computationally intensive. Maybe that would help with the long wait times to run your analysis script?

### Summary assessment
* well structured

## Documentation 
How well is the project documented? Are you able to understand each step of the whole analysis, each decision that was made, and each line of code? Is enough information provided as comments in code or as part of Rmd files?

### Feedback and Comments

Basically see above. It was really easy to follow along with what you did an discern how and why you did things from text in the code files themselves. No recommendations.

### Summary assessment
* fully and well documented

## Reproducibility
Are all results fully reproducible? Is documentation provided which clearly explains how to reproduce things, and does it work without the need for any manual intervention? Are you able to re-run the whole analysis without having to do manual interventions/edits?

### Feedback and Comments

Both your processing script and statistical analysis script produced errors and I was not able to reproduce your results. In your processing script, it was a problem with your "xlsx" package. I wan not able to even install it, so I couldnt load the data. In the statistical analysis script, I got this error on the last chunk: file_location <- here("results", "Figures", "ROC_TestPlot").ggsave(filename = file_location, plot = ROC_Test_RF)
Error: Unknown graphics device ''. I am not sure what this means, but I would check on it.

### Summary assessment
* major parts not reproducible

## Thoroughness
How thorough was the overall study? Were alternatives (e.g. different ways of processing the data or different models) considered? Were alternatives discussed? Were the questions/hypotheses fully and thoroughly addressed?

### Feedback and Comments

Based on diversity of exploration and analytical tests performed, I was impressed with the thoroughness of your different analyses. No recommendations

### Summary assessment
* strong level of thoroughness

## Further comments

_The only glaring issue is that portions of your processing script and statistical analysis script did not run. Just check the issues described above and I think you are all set! Overall, very well done!_

### 2021-12-06 products/manuscript/Manuscript.Rmd
Peer Review Edits to the MS
Peer review edits from the 2 reviewers. Almost completed for submission.

@@
 
-## Abstract
-
-Blood parasites of the *Haemoproteus* genus can infect various bird species and cause chronic infections that last an entire lifetime.
-A general increase in the presence of parasitic pathogens of birds led to the collection of blood samples from American White Ibis in Southern Florida from 2010-2017.
-A data analysis on the demographic, temporal, spatial, and parasitemia data collected from over 300 sampled Ibis was conducted.
-Various univariate and multivariate logistic regression models, along with a random forest model, concluded that the sampling date was the most important predictor in determining whether a sampled Ibis had a *Haemoproteus* infection (ROC-AUC 0.725, Accuracy 0.715).
-This led to the observation that more Ibis have *Haemoproteus* infections in Southern Florida in 2017 compared to 2010, however, there are numerous other epidemiologic and pathogenic factors about the parasite that makes any further conclusions difficult.
-An increased sampling effort and further studies are needed.
-
-## Loading Required Data and Packages
-```{r}
+```{r, echo = FALSE, message = FALSE}
 #Loading required packages #for data processing
@@
 
-## Introduction
+# Abstract
+
+Blood parasites of the *Haemoproteus* genus can infect various bird species and cause chronic infections that last an entire lifetime.
+A general increase in the presence of parasitic pathogens of birds led to the collection of blood samples from American White Ibis in Southern Florida from 2010-2017.
+A data analysis on the demographic, temporal, spatial, and parasitemia data collected from over 300 sampled Ibis was conducted.
+Various univariate and multivariate logistic regression models, along with a random forest model, concluded that the sampling date was the most important predictor in determining whether a sampled Ibis had a *Haemoproteus* infection (ROC-AUC 0.725, Accuracy 0.715).
+This led to the observation that more Ibis have *Haemoproteus* infections in Southern Florida in 2017 compared to 2010, however, there are numerous other epidemiologic and pathogenic factors about the parasite that makes any further conclusions difficult.
+An increased sampling effort and further studies are needed.
+
+# Introduction
 
@@
 
-### General Background Information
+## General Background Information
 The data this analysis concerns is American White Ibis (*Eudocimus albus*) blood parasite data that was collected from 2010-2017.
@@
 Lower parasitemia values suggest that the infections in the Ibis are chronic [@Coker2017].
+This is a risk for the White Ibis population to be impacted, as well as multiple other host species, and these parasites could have a negative population level impact on them even if they don't cause acute mortality.
+Analyzing this data will help identify any demographic or spatiotemporal trends among infected Ibis in order to help possibly assist management of the species and parasite in Southern Florida.
 
@@
 Here we investigate whether age, sex, sampling site, capture season, date of capture, body condition, and mass of captured Ibis make significant differences on the overall presence of *Haemoproteus* in the birds.
+We expect *Haemoproteus* presence to change between capture season and date, as well as body condition and overall mass of the birds.
 
-## Methods
+# Methods
 
-### Data Collection
+## Data Collection
 
@@
 
-### Data Cleaning
+## Data Cleaning
 
@@
 
-### Data Analysis
+## Data Analysis
 
@@
 
-### Study Population
+## Study Population
 
@@
 
-### Exploratory Analysis
+## Exploratory Analysis
 
@@
 
-#### Tables
-```{r}
+### Tables
+```{r, echo = FALSE, message = FALSE}
 IbisCount1
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 IbisCount2
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 IbisCount3
@@
 
-#### Figures
+### Figures
 
@@
 Figure 1: Haemoproteus parasitemia from sampled infected Ibises from 2010-2017 by habitat type
-```{r}
+```{r, echo = FALSE, message = FALSE}
 knitr::include_graphics(figure_Exp_1, error = FALSE)
@@
 Figure 2: Haemoproteus parasitemia by total female and male Ibis mass
-```{r}
+```{r, echo = FALSE, message = FALSE}
 knitr::include_graphics(figure_Exp_4, error = FALSE)
@@
 Figure 3: Haemoproteus parasitemia by sample site and habitat type
-```{r}
+```{r, echo = FALSE, message = FALSE}
 knitr::include_graphics(figure_Exp_6, error = FALSE)
@@
 
-### Statistical Analyses
+## Statistical Analyses
 
-#### Univariate and Multivariate Logistic Models
+### Univariate and Multivariate Logistic Models
 
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 model4
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 model5
@@
 
-```{r}
+```{r, echo = FALSE, message = FALSE}
 model8
@@
 
-### Random Forest Model
+## Random Forest Model
 
 A random forest model was run on all the predictors against the *Haemoproteus* parasitemia presence and absence data.
+This model was the best of 3 machine learning models tested including a decision tree and LASSO model (refer to supplemental materials folder).
 The number of features to consider at each split and the minimum number of data points required for the node to be split were both tuned with a tuning grid.
@@
 
-```{r Variable Importance Model}
+```{r Variable Importance Model, echo = FALSE, message = FALSE}
 knitr::include_graphics(figure_ML_4, error = FALSE)
@@
 
-```{r ROC Curve for Random Forest Model, echo = FALSE, out.width = '100%'}
+```{r ROC Curve for Random Forest Model, echo = FALSE, message = FALSE, out.width = '100%'}
 knitr::include_graphics(figure_ML_5, error = FALSE)
@@
 
-### Discussion
+## Discussion
 
@@
 The other individual predictors were all used in univariate logistic regression models, and none of them had highly significant results.
+These results suggest that much further research is needed, potentially continuing a multi-year study on the parasite that could provide stronger evidence for the relationaships between the variables analyzed here.
+Potentially looking at *Haemoproteus* presence in *Culicoides* midges in the area could be helpful as well.
 
-### Conclusions
+## Conclusions
 
@@
 However, there are numerous other factors regarding the *Haemoproteus* epidemiology and pathogenesis that make any further conclusions difficult.
+
+## References

### 2021-12-06 analysis_scripts/StatisticalAnalysisModels.R

@@ -430,6 +430,9 @@ rf_predobs_plot <- ggplot(rf_residpredict, aes(x = HaeParasitPA, y = .pred_1)) +
   ggtitle("Predicted vs. Observed Values for Random Forest Model")
 rf_predobs_plot
 
+#The RANDOM FOREST MODEL performed the best on the training dataset out of all
+#the models tested.
+
 ##############################Fitting Random Forest to Test Data########################################
 set.seed(144)
 Test_rf_fit <- final_rf_wf %>% last_fit(data_split_log)

### 2021-12-06 processing_script/ProcessingScript.R

@@ -3,6 +3,7 @@
 #This script loads and cleans raw data for the Ibis data analysis process. It then saves processed data for further analysis.
 
 ###############################Loading all required packages###########################
+library(rJava) #Required for xlsx. Must have Java installed and installed this package.
 library(here) #to set paths
 library(tidyverse) #all required data manipulation packages
 library(xlsx) #to read in excel spreadsheets
